---
title: "Analyzing Git Log Entities in Parallel"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Analyzing Git Log Entities in Parallel}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This notebook showcases the git log entity analysis over multiple time windows.

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
Sys.setlocale("LC_ALL","C") # set default locale
```

To speed up the analysis, the time windows are processed in parallel in a foreach-loop. For this purpose, the following packages are required:

* doParallel (>= 1.0.17)
* doSNOW (>= 1.0.20)

```{r warning=FALSE,message=FALSE}
require(data.table)
require(doParallel)
require(doSNOW)
require(igraph)
require(kaiaulu)
require(knitr)
require(magrittr)
require(reactable)
require(stringi)
require(visNetwork)
require(yaml)
```

# Project Configuration File

```{r}
tool <- yaml::read_yaml("../tools.yml")
conf <- yaml::read_yaml("../conf/spark.yml")

# 3rd Party Tools
perceval_path <- tool[["perceval"]]
utags_path <- tool[["utags"]]

# Ctags Line Types
kinds <- conf[["tool"]][["uctags"]][["keep_lines_type"]]

# Local Git Repo Folder Path
git_repo_path <- conf[["version_control"]][["log"]]

# File Filters
file_extensions <- conf[["filter"]][["keep_filepaths_ending_with"]]
substring_filepath <- conf[["filter"]][["remove_filepaths_containing"]]

# Analysis Parameters
window_size <- conf[["analysis"]][["window"]][["size_days"]]

# Results Save Path
save_path <- "../../results/spark"
```

# Construct Collaboration Network

```{r}
gitlog_path <- file.path(save_path, "gitlog.rds")
project_git <- parse_gitlog(perceval_path,git_repo_path,gitlog_path)
#project_git <- readRDS(gitlog_path)
```

# Filter files

```{r}
project_git <- project_git  %>%
  filter_by_file_extension(file_extensions,"file_pathname")  %>% 
  filter_by_filepath_substring(substring_filepath,"file_pathname")
```

```{r}
project_git$author_datetimetz <- as.POSIXct(project_git$author_datetimetz,
                                              format = "%a %b %d %H:%M:%S %Y %z", tz = "UTC")
project_git$author_datetimetz <- as.POSIXct(project_git$committer_datetimetz,
                                              format = "%a %b %d %H:%M:%S %Y %z", tz = "UTC")
```

# Multi-Core Cluster Setup

Set up the parallel computing cluster to parallelize foreach-loops for time window analysis.

```{r}
# Configure the cluster
n.cores <- parallel::detectCores() # Use all available cores
cluster <- parallel::makeCluster(n.cores, outfile="")
registerDoSNOW(cluster) # Register the cluster for foreach loops

paste("Registered workers for foreach loops:", foreach::getDoParWorkers())
```

# Parallel Time Window Entity Analysis

Entity analysis requires a much longer time to process than file analysis, which is why the associated function provides a progress bar. Since multiple time windows are analyzed in parallel, another progress bar is stacked on the individual ones to indicate the overall status. 

Please note that the following cell analyzes the entire git history, which may take several minutes to hours.

```{r}
# The start and end dates correspond to the date of the earliest and latest commit, respectively.
start_date <- min(project_git$author_datetimetz,na.rm=TRUE)
end_date <- max(project_git$author_datetimetz,na.rm=TRUE)

# Define all timestamps in number of days since the very first commit of the repo
# Format time window for posixT
window_size_f <- stringi::stri_c(window_size," day")

# Note if end_date is not (and will likely not be) a multiple of window_size, 
# then the ending incomplete window is discarded so the metrics are not calculated 
# in a smaller interval
time_window <- seq.POSIXt(from=start_date,to=end_date,by=window_size_f)
size_time_window <- length(time_window)

# Progress bar
pb <- txtProgressBar(min=0, max=size_time_window-1, style=3, char=" ", width=85)
progress <- function(n) setTxtProgressBar(pb, n)
opts <- list(progress = progress)

k <- 0
foreach(
  j = 2:size_time_window,
  .packages = c("kaiaulu", "reactable", "data.table", "stringi"),
  .options.snow = opts
) %dopar% {
  i <- j - 1
  k <- k + 1

  # If the time window is of size 1, then there have been less than
  # "window_size_f" days from the start date.
  if(length(time_window) == 1){
    start_day <- start_date
    end_day <- end_date
  }else{
    start_day <- time_window[i]
    end_day <- time_window[j]
  }

  # Define the save path for the entity log
  start_day_formatted <- format(start_day, "%Y-%m-%d")
  end_day_formatted <- format(end_day, "%Y-%m-%d")
  window_save_path = file.path(save_path, "entity", paste(start_day_formatted, "_", end_day_formatted, sep=""))
  entity_save_path <- file.path(window_save_path, "gitlog_entity.rds")

  # Obtain all commits from the gitlog which are within a particular window_size
  project_git_slice <- project_git[(project_git$author_datetimetz >= start_day) &
                                     (project_git$author_datetimetz < end_day), ]

  # Perform entity analysis
  if (nrow(project_git_slice) > 0) {
    changed_entities <- parse_gitlog_entity(git_repo_path,utags_path,project_git_slice,kinds,progress_bar=TRUE)
    
    if (nrow(changed_entities)>0) {
      # Naming
      changed_entities[,c("author_name_email",
                          "author_datetimetz",
                          "committer_name_email",
                          "commit_hash",
                          "committer_datetimetz",
                          "entity",
                          "weight"):=list(stri_c(changed_entities$author_name,changed_entities$author_email,sep= " "),
                                          as.POSIXct(as.integer(changed_entities$author_timestamp),
                                                     origin="1970-01-01",tz = "UTC"),
                                          stri_c(changed_entities$committer_name,changed_entities$committer_email,sep= " "),
                                          changed_entities$commit_hash,
                                          as.POSIXct(as.integer(changed_entities$committer_timestamp),
                                                     origin="1970-01-01",tz = "UTC"),
                                          changed_entities$entity_definition_name,
                                          changed_entities$n_lines_changed
                                          )]
      
      # Identity match
      project_log <- list(project_git=changed_entities)
      project_log <- identity_match(project_log,
                                    name_column = c("author_name_email"),
                                    assign_exact_identity,
                                    use_name_only=TRUE,
                                    label = "raw_name"
                                    )
      window_log <- project_log[["project_git"]]
      
      # Save results
      dir.create(window_save_path, recursive = TRUE)
      saveRDS(window_log, file = entity_save_path)
    }
  }
}
```

# Parallel Clean-Up

Free computing resources after parallelized computations.

```{r}
close(pb)
stopCluster(cluster)
```

# Temporal Networks 

The results stored for each time window may be used for arbitrary analyses. For example, they can be used to inspect the developers' collaboration from a temporal perspective.

```{r}
# Load the entities for the time window to inspect
window_path <- file.path(save_path, "entity", "2013-03-13_2013-12-08", "gitlog_entity.rds")
git_log_entities <- readRDS(window_path)

# Plot the temporal projection
project_temporal_collaboration_network <- transform_gitlog_to_entity_temporal_network(git_log_entities)
project_temporal_collaboration_network <- igraph::graph_from_data_frame(d=project_temporal_collaboration_network[["edgelist"]],
                      directed = TRUE,
                      vertices = project_temporal_collaboration_network[["nodes"]])
visIgraph(project_temporal_collaboration_network,randomSeed = 1)
```

