---
title: "Lift -- Aiwork Model from Git Commit Subjects"
output:
  html_document:
    toc: true
    number_sections: true
params:
  commits: "commits.csv"   # input: extract output (kaiaulu::parse_gitlog)
  out:     "aiwork.csv"    # output: model-language CSV (one row per project)
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Aiwork Churn Velocity}
  %\VignetteEncoding{UTF-8}
---

```{r}
set.seed(1)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu)
require(data.table)
require(stringi)
```

# Introduction

The `aiwork` SD model (GitClear 2024 + METR 2025) carries five
parameters; two of them can be lifted directly from commit history
without per-developer AI-usage telemetry:

* **`churn_base`** -- baseline (no-AI) churn rate.  Estimated as the
  fraction of commits whose subject matches a revert / rollback /
  "fix typo" pattern.  GitClear and METR both treat AI-induced
  churn as a multiplicative inflation *on top of* this base.

* **`mature_rate`** -- the Wip -> Kept transition rate.  Lifted as
  the inverse of the mean short-author commit-cycle length (days
  between the same author's first and last commit in a short run).
  A rough proxy for "how fast WIP becomes shipped" pending PR-cycle
  data from issue trackers.

The remaining params (`gen_boost`, `churn_mult`, `verify_drag`)
remain literature priors -- they require AI-usage telemetry the
OSS git log cannot supply.

# Method

1.  Read the extract-stage commits CSV (already produced by
    `kaiaulu::parse_gitlog`).
2.  Match each commit subject against the churn regex.
3.  `churn_base` = matched / total commits, capped to [0, 1].
4.  Per-author short-run cycle: max(date) - min(date) within author;
    `mature_rate` = 1 / median(cycle in ticks), where 1 tick = 7 days.
5.  Write a one-row CSV with the two lifted values + diagnostic counts.

# Lift

```{r}
commits <- fread(params$commits)
setnames(commits, tolower(names(commits)))

# kaiaulu schema: author_name_email, author_datetimetz, commit_hash, commit_message, ...
setnames(commits, "author_name_email", "author")
# Dedupe by commit (the CSV has one row per commit-file).
commits <- unique(commits, by = "commit_hash")

# Churn-subject pattern: revert/rollback/fix-typo/undo conventions.
churn_rx <- "(?i)\\b(revert|rollback|undo|fix(:|\\s)?\\s*typo|hot ?fix|re-?do)\\b"
matched  <- stri_detect_regex(commits$commit_message, churn_rx)
churn_base <- min(1, max(0, mean(matched, na.rm = TRUE)))

# mature_rate: inverse of median author-run length, expressed in ticks.
commits[, date := as.IDate(as.POSIXct(author_datetimetz,
                                      format = "%a %b %d %H:%M:%S %Y %z",
                                      tz = "UTC"))]
runs <- commits[!is.na(date), .(span_days = as.integer(max(date) - min(date)) + 1),
                by = author]
runs <- runs[span_days > 0]
med_days  <- median(runs$span_days)
ticks     <- med_days / 7
mature_rate <- if (ticks > 0) 1 / ticks else NA_real_

# Output: one row, the two lifted aiwork params + provenance counts.
out <- data.table(
  metric        = c("churn_base", "mature_rate"),
  value         = c(churn_base, mature_rate),
  n_commits     = nrow(commits),
  n_churn_match = sum(matched, na.rm = TRUE),
  n_authors     = nrow(runs))
fwrite(out, params$out)
out
```

# Notes

* No PR-cycle data => `mature_rate` is a proxy.  When issue-tracker
  extracts are available, swap in the PR-merge cycle directly.
* `churn_base` regex is conservative.  Project-specific conventions
  (e.g., "amend", "squash") could widen it; documented here as a
  honest under-count.
