---
title: "Lift — Archpat Model from Apache Helix (GoF + RefactoringMiner)"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Archpat Gof Pattern Partition}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu); require(data.table); require(stringi)
require(knitr); require(yaml); require(magrittr); require(jsonlite)
# Helpers (myths_*.R) loaded with library(kaiaulu).
```

# Introduction

The `archpat` model (Ric Kazman's claim) tests whether design
patterns applied to an already-bad codebase can pay down architectural
debt. The SD form (`models/sd.py:archpat()`) has three architectural
stocks plus a Debt accumulator:

```
Legacy ──migrate──> Patterned ──decay_rate──> Drift ──drift_to_legacy──> Legacy
            (effort applied)     (Perry-Wolf erosion)
```

`ctrl` = `migrate`. Thesis: starting from
`Patterned=10, Legacy=90, Debt=40`, aggressive migration
(`migrate=1.5`) repairs the project faster than slow migration
(`migrate=0.2`).

**Method outline** (full chain):

1. `mvn compile -pl helix-core -am -DskipTests` — produces `.class`
   files in `target/classes/`.
2. `java -jar pattern4.jar -target <classes> -output <xml>` — GoF
   patterns per class. Discovered today: pattern4 IS CLI-callable
   despite its GUI-default manifest (the dispatch is in
   `MatrixFrame.main`). See memory:
   `reference_pattern4_gotcha.md` for full incantation.
3. `parse_pattern4_xml.py <pattern4-dir>` — Python parser, emits a
   `patterned_files.csv` with `file_pathname, pattern_type, role,
   module` columns. (We do this in Python because R's kaiaulu wrapper
   `parse_gof_patterns()` is a parser, not a runner — same data, two
   languages.)
4. Bug-frequency signal: SZZ-introducing-commits touching each file
   (`scripts/szz_helix.py` → `szz_pairs.csv`). Replaces JIRA-Bug
   filter because the partial JIRA dump on disk has 0 Bug-type
   issues.
5. File-churn: `compute_file_churn()` from `functions.R`.
6. Assign each .java file to one of {Patterned, Legacy, Drift, Other}
   via `assign_file_partition()`.

**Fallback path**: if pattern4 setup were unavailable, Arcan
(structural smells) is the kaiaulu-documented substitute. Arcan
output has the same `(file, smell_type, smell_id)` shape; the
partition logic is detector-agnostic. Document the substitution in
any paper deliverable since smells ≠ GoF patterns semantically.

# Configuration

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/helix.yml")

perceval_path <- get_tool_project("perceval", tool)
git_repo_path <- get_git_repo_path(conf)
```

# Parse Git Log

```{r}
project_git <- parse_gitlog(perceval_path, git_repo_path) %>%
  filter_by_file_extension(get_file_extensions(conf), "file_pathname") %>%
  filter_by_filepath_substring(get_substring_filepath(conf),
                               "file_pathname")
project_git$author_datetimetz <- as.POSIXct(project_git$author_datetimetz,
                                            format = "%a %b %d %H:%M:%S %Y %z",
                                            tz = "UTC")
cat("gitlog rows:", nrow(project_git),
    " commits:", uniqueN(project_git$commit_hash), "\n")
```

# Patterned Partition (pattern4)

The pattern4 + Python parser run is OUT-OF-NOTEBOOK (pre-knit) since
`mvn compile` is a multi-minute side effect. The produced CSV lives
at `../data/helix/derived/pattern4/patterned_files.csv`.

```{r}
pat_dt <- fread("../data/helix/derived/pattern4/patterned_files.csv")
patterned_files <- unique(pat_dt$file_pathname)
cat("Modules scanned:", paste(unique(pat_dt$module), collapse = ", "), "\n")
cat("Pattern instances:", nrow(pat_dt),
    "  unique patterned files:", length(patterned_files), "\n")
cat("Pattern type distribution:\n")
print(pat_dt[, .N, by = pattern_type][order(-N)])
```

# Bug-Frequency Signal (SZZ)

```{r}
szz <- parse_szz_bugfixes("../data/helix/derived/szz_pairs.csv")
cat("SZZ pairs:", nrow(szz),
    "  unique introducing commits:", uniqueN(szz$introducing_commit_hash), "\n")

intro_hashes <- unique(szz$introducing_commit_hash)
intro_files  <- unique(project_git[commit_hash %in% intro_hashes,
                                   .(commit_hash, file_pathname)])
file_bug_freq <- intro_files[, .(bug_count = uniqueN(commit_hash)),
                              by = file_pathname]
nrow(file_bug_freq)
```

# Churn Signal

```{r}
file_churn <- compute_file_churn(project_git, window_days = 180)
nrow(file_churn)
```

# Assign Partition

```{r}
partition <- assign_file_partition(
  patterned_files       = patterned_files,
  file_bug_freq         = file_bug_freq,
  file_churn            = file_churn,
  legacy_bug_threshold  = 3,
  drift_churn_threshold = 0.5
)
print(table(partition$stock, useNA = "ifany"))
```

# Outputs for the SD Framework

```{r}
counts <- table(partition$stock)
out <- data.table(
  project           = "helix",
  Patterned_n       = as.integer(counts["Patterned"]),
  Legacy_n          = as.integer(counts["Legacy"]),
  Drift_n           = as.integer(counts["Drift"]),
  Other_n           = as.integer(counts["Other"]),
  n_files           = nrow(partition),
  n_patterned_input = length(patterned_files),
  modules           = paste(unique(pat_dt$module), collapse = ","),
  seed              = seed
)
fwrite(out, "../outputs/lift_archpat_helix.csv")
kable(out)
```

# Verdict on Helix

Helix's partition: ~6.5% Patterned (149 files), ~19% Legacy
(384 files), 0 Drift (no file passes the 0.5 churn threshold in the
recent 180-day window — Helix is a mature, stable codebase), ~74%
Other.

Running `models/sd.py:archpat.rq(bg=helix_calib)` with these stocks
(plugged in via `scripts/calibrate.py`) widens the thesis gap from
+229 (default) to +390. **Helix's bigger Legacy stock strengthens
the repair thesis** — more legacy means more headroom for migrate to
move files into the Patterned bucket.

**Family-member test on Ambari** (2nd Java + Maven project):

| project | Patterned | Legacy   | Drift | Other  | n_files |
|---------|----------:|---------:|------:|-------:|--------:|
| Helix   | 149       | 384 (OUT)| 0     | 1,452  | 1,985   |
| Ambari  | 381 (OUT) | 1,890 (OUT) | 0  | 4,329  | 6,600   |

**Boundary-adequacy failure**: archpat's `Legacy` stock has model
declared `hi = 200`. Both Helix (384) and Ambari (1,890) exceed.
`Patterned hi = 200` also exceeded by Ambari (381). The model's
bounds were specified at small-project scale and don't span mature
OSS. Recommend widening to ≥ 1000 / ≥ 3000 respectively. See
`findings.md` F0 for the full boundary-violation table.

# Sanity Checks

**(1) Bug-count dependency**: this lift requires bug-classification
per commit. We use SZZ-introducing-commit-touches as a proxy because
the partial JIRA dump on disk has 0 Bug-type issues. With the full
Carlos-cleaned Helix JIRA dump (on his Drive), we could tighten this
to JIRA-Bug-type-filtered commits, narrowing false-positives.

**(2) Identity bridging**: this lift uses git only, not comms. No
cross-source identity_match required.

# Pattern4 Setup Notes (memory: reference_pattern4_gotcha.md)

The pattern4.jar from
https://users.encs.concordia.ca/~nikolaos/files/pattern_detection/pattern4.jar
declares `Main-Class: gr.uom.java.pattern.gui.MatrixFrame`, which
*opens a Swing GUI* on `java -jar pattern4.jar` with no args. But
`MatrixFrame.main(args)` dispatches:

```java
if (args.length == 4) {
  if (args[0].equals("-target") && args[2].equals("-output")) {
    new Console(inputDir, outputXML);  // batch mode!
  }
} else {
  new MatrixFrame();  // GUI
}
```

So:
```bash
java -Xmx2g -jar pattern4.jar -target <classes-dir> -output <xml>
```
runs the batch detector. This was NOT obvious until we read the
decompiled main; an earlier session memory had "pattern4 is GUI-only"
which is wrong. Corrected in
`~/.claude/projects/.../memory/reference_pattern4_gotcha.md`.

# References

- Martin, R. C. (2008). *Clean Architecture*.
- Perry, D. E. & Wolf, A. L. (1992). Foundations for the study of
  software architecture. *ACM SIGSOFT Software Engineering Notes*.
- Tsantalis, N. et al. — pattern4 GoF detector
  (https://users.encs.concordia.ca/~nikolaos/pattern_detection.html).
- Arcelli Fontana, F. et al. — Arcan smell detector
  (https://essere.disco.unimib.it/wiki/arcan/) — documented fallback.
- `models/sd.py:archpat()` — the SD model.
- `findings.md` F0 — multi-project boundary-violation summary.
