---
title: "Lift -- Archpat from Pre-Extracted Commits + LoC + Patterns"
output:
  html_document:
    toc: true
    number_sections: true
params:
  commits:  "commits.csv"
  loc:      "loc.csv"
  patterns: "patterns.csv"
  out:      "archpat.csv"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Archpat Lift From Commits}
  %\VignetteEncoding{UTF-8}
---

```{r}
set.seed(1)
```

```{r warning=FALSE, message=FALSE}
require(data.table)
require(stringi)
```

# Method

Consumes three pre-extracted CSVs and produces the archpat-model
inputs:

* `commits.csv` (kaiaulu::parse_gitlog) -> commit volume per module
* `loc.csv` (scc) -> module sizes (loc per file)
* `patterns.csv` (pattern4.jar) -> GoF pattern occurrences per file

Outputs four lift values:

* `Patterned_n`  -- module count under a detected GoF pattern
* `Legacy_n`     -- remaining module count
* `gen_pat_rate` -- mean commits/patterned-module/month (proxy for
                    archpat's `gen_pat` parameter)
* `gen_leg_rate` -- same for non-patterned modules (`gen_leg` proxy)

# Lift

```{r}
commits  <- fread(params$commits)
loc      <- fread(params$loc)
patterns <- fread(params$patterns)
setnames(commits,  tolower(names(commits)))
setnames(loc,      tolower(names(loc)))
setnames(patterns, tolower(names(patterns)))

# Patterned modules = unique file/class paths flagged by pattern4.
patterned_files <- unique(patterns$file)

# All known modules (from scc LoC dump): one row per file.
all_files <- unique(loc$location)
patterned_set <- intersect(all_files, patterned_files)
legacy_set    <- setdiff(all_files, patterned_files)

Patterned_n <- length(patterned_set)
Legacy_n    <- length(legacy_set)

# kaiaulu schema: one row per commit-file with file_pathname + author_datetimetz.
file_col <- "file_pathname"
commits[, date := as.POSIXct(author_datetimetz,
                             format = "%a %b %d %H:%M:%S %Y %z",
                             tz = "UTC")]
commits <- commits[!is.na(date) & !is.na(get(file_col))]
months <- as.numeric(difftime(max(commits$date), min(commits$date), units = "days")) / 30
months <- max(1, months)

pat_commits <- nrow(commits[get(file_col) %in% patterned_set])
leg_commits <- nrow(commits[get(file_col) %in% legacy_set])

gen_pat_rate <- if (Patterned_n > 0) pat_commits / Patterned_n / months else NA_real_
gen_leg_rate <- if (Legacy_n    > 0) leg_commits / Legacy_n    / months else NA_real_

out <- data.table(
  metric = c("Patterned_n", "Legacy_n", "gen_pat_rate", "gen_leg_rate"),
  value  = c(Patterned_n,    Legacy_n,   gen_pat_rate,   gen_leg_rate),
  n_pat_commits = pat_commits,
  n_leg_commits = leg_commits,
  months_span   = months)
fwrite(out, params$out)
out
```

# Notes

* Pattern4 detects GoF patterns in Java only. For polyglot repos, the
  Legacy stock will be inflated.
* gen_*_rate uses module *count* as denominator (not LoC); swap to
  per-loc rate for a more even comparison.
