---
title: "Lift — Archpat Model from Apache Helix (GoF + RefactoringMiner)"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Archpat Gof Pattern Partition}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu); require(data.table); require(stringi)
require(knitr); require(yaml); require(magrittr); require(jsonlite)
# Helpers (myths_*.R) loaded with library(kaiaulu).
```

# Introduction

This notebook walks the GoF-pattern partition lift used to calibrate
the `archpat` system-dynamics model. The model tests the
software-engineering claim that design patterns applied to an
already-degraded codebase can pay down architectural debt — the
Patterned vs Legacy code partition. The claim traces to:

- Gamma, Helm, Johnson, Vlissides (1994). *Design Patterns: Elements
  of Reusable Object-Oriented Software*. Addison-Wesley.
- Tsantalis et al. (2006). "Design Pattern Detection Using Similarity
  Scoring." *IEEE Trans. Software Engineering* 32(11):896–909
  (pattern4.jar's algorithmic basis).
- MacCormack, Rusnak, Baldwin (2007). "Exploring the Structure of
  Complex Software Designs: An Empirical Study of Open Source and
  Proprietary Code." *Management Science* 52(7):1015–1030.

The SD form encodes three architectural stocks plus a Debt
accumulator:

```
Legacy ──migrate──> Patterned ──decay_rate──> Drift ──drift_to_legacy──> Legacy
            (effort applied)     (Perry-Wolf erosion)
```

The thesis is that starting from `Patterned=10, Legacy=90, Debt=40`,
aggressive migration (`migrate=1.5`) repairs the project faster than
slow migration (`migrate=0.2`).

**Method outline** (full chain):

1. `mvn compile -pl helix-core -am -DskipTests` — produces `.class`
   files in `target/classes/`. Multi-minute side effect; run
   externally before knitting.
2. `java -jar pattern4.jar -target <classes> -output <xml>` —
   detects 23 GoF patterns per class. pattern4 is CLI-callable
   despite its GUI-default manifest: `MatrixFrame.main(args)` parses
   `-target` and `-output` switches before opening the Swing view.
3. `parse_pattern4_xml.py <pattern4-dir>` — Python parser; emits
   `patterned_files.csv` with `file_pathname, pattern_type, role,
   module` columns. R-side kaiaulu offers
   `\code{\link{parse_gof_patterns}}` for reading existing XML
   output; the Python parser is used here only as a runner.
4. **Bug-frequency signal**: SZZ-introducing-commits touching each
   file (an external script produces `szz_pairs.csv`). The standard
   `\code{\link{parse_jira}}` Bug-type filter is the kaiaulu-native
   alternative when a complete JIRA dump is available.
5. **File-churn**: `\code{\link{compute_file_churn}}`.
6. Assign each `.java` file to one of `{Patterned, Legacy, Drift,
   Other}` via `\code{\link{assign_file_partition}}`.

**Fallback path**: where pattern4 is unavailable, Arcan (Arcelli
Fontana, Pigazzini, Roveda, Tamburri, Zanoni, Lenarduzzi 2017 —
"Arcan: A tool for architectural smells detection," *ECSA 2017*)
produces structural-smell output with the same
`(file, smell_type, smell_id)` shape; the partition logic is
detector-agnostic. Smells &ne; GoF patterns semantically — document
the substitution in any downstream comparison.

# Configuration

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/helix.yml")

perceval_path <- get_tool_project("perceval", tool)
git_repo_path <- get_git_repo_path(conf)
```

# Parse Git Log

```{r}
project_git <- parse_gitlog(perceval_path, git_repo_path) %>%
  filter_by_file_extension(get_file_extensions(conf), "file_pathname") %>%
  filter_by_filepath_substring(get_substring_filepath(conf),
                               "file_pathname")
project_git$author_datetimetz <- as.POSIXct(project_git$author_datetimetz,
                                            format = "%a %b %d %H:%M:%S %Y %z",
                                            tz = "UTC")
cat("gitlog rows:", nrow(project_git),
    " commits:", uniqueN(project_git$commit_hash), "\n")
```

# Patterned Partition (pattern4)

The pattern4 + Python parser run is OUT-OF-NOTEBOOK (pre-knit) since
`mvn compile` is a multi-minute side effect. The produced CSV lives
at `../data/helix/derived/pattern4/patterned_files.csv`.

```{r}
pat_dt <- fread("../data/helix/derived/pattern4/patterned_files.csv")
patterned_files <- unique(pat_dt$file_pathname)
cat("Modules scanned:", paste(unique(pat_dt$module), collapse = ", "), "\n")
cat("Pattern instances:", nrow(pat_dt),
    "  unique patterned files:", length(patterned_files), "\n")
cat("Pattern type distribution:\n")
print(pat_dt[, .N, by = pattern_type][order(-N)])
```

# Bug-Frequency Signal (SZZ)

```{r}
szz <- parse_szz_bugfixes("../data/helix/derived/szz_pairs.csv")
cat("SZZ pairs:", nrow(szz),
    "  unique introducing commits:", uniqueN(szz$introducing_commit_hash), "\n")

intro_hashes <- unique(szz$introducing_commit_hash)
intro_files  <- unique(project_git[commit_hash %in% intro_hashes,
                                   .(commit_hash, file_pathname)])
file_bug_freq <- intro_files[, .(bug_count = uniqueN(commit_hash)),
                              by = file_pathname]
nrow(file_bug_freq)
```

# Churn Signal

```{r}
file_churn <- compute_file_churn(project_git, window_days = 180)
nrow(file_churn)
```

# Assign Partition

```{r}
partition <- assign_file_partition(
  patterned_files       = patterned_files,
  file_bug_freq         = file_bug_freq,
  file_churn            = file_churn,
  legacy_bug_threshold  = 3,
  drift_churn_threshold = 0.5
)
print(table(partition$stock, useNA = "ifany"))
```

# Outputs for the SD Framework

```{r}
counts <- table(partition$stock)
out <- data.table(
  project           = "helix",
  Patterned_n       = as.integer(counts["Patterned"]),
  Legacy_n          = as.integer(counts["Legacy"]),
  Drift_n           = as.integer(counts["Drift"]),
  Other_n           = as.integer(counts["Other"]),
  n_files           = nrow(partition),
  n_patterned_input = length(patterned_files),
  modules           = paste(unique(pat_dt$module), collapse = ","),
  seed              = seed
)
fwrite(out, "../outputs/lift_archpat_helix.csv")
kable(out)
```

# Verdict on Helix

Helix's partition: ~6.5% Patterned (149 files), ~19% Legacy
(384 files), 0 Drift (no file passes the 0.5 churn threshold in the
recent 180-day window — Helix is a mature, stable codebase), ~74%
Other.

Helix's bigger Legacy stock means more headroom for the `migrate`
control to move files into the Patterned bucket; a downstream SD
verdict pipeline that consumes these stocks therefore reports a
larger thesis-gap than its defaults predict.

**Family-member comparison on Ambari** (2nd Java + Maven project):

| project | Patterned | Legacy   | Drift | Other  | n_files |
|---------|----------:|---------:|------:|-------:|--------:|
| Helix   | 149       | 384      | 0     | 1,452  | 1,985   |
| Ambari  | 381       | 1,890    | 0     | 4,329  | 6,600   |

Both projects produce Legacy counts well above 200; for downstream
SD models with a Legacy-stock upper bound near that magnitude, the
empirical counts will trigger boundary-clipping during simulation.

# Sanity Checks

**(1) Bug-count dependency**: this lift requires per-commit defect
classification. We use SZZ-introducing-commit-touches because the
JIRA dump available here lacks Bug-type issues. With a full JIRA
dump, swap in `\code{\link{parse_jira}}` and filter on
`issuetype == "Bug"` for tighter false-positive control.

**(2) Identity bridging**: this lift uses git only, not comms. No
cross-source `\code{\link{identity_match}}` required.

# Pattern4 Invocation

The pattern4.jar declares `Main-Class: gr.uom.java.pattern.gui.MatrixFrame`
which opens a Swing GUI on `java -jar pattern4.jar` with no args.
`MatrixFrame.main(args)` dispatches to a batch-mode `Console` when
called as:

```bash
java -Xmx2g -jar pattern4.jar -target <classes-dir> -output <xml>
```

The GUI path is taken only when no recognised switches appear in
`args`. The batch invocation is the one this vignette assumes.

# References

- Gamma, E., Helm, R., Johnson, R., & Vlissides, J. (1994).
  *Design Patterns: Elements of Reusable Object-Oriented Software*.
  Addison-Wesley.
- Perry, D. E. & Wolf, A. L. (1992). Foundations for the study of
  software architecture. *ACM SIGSOFT Software Engineering Notes*
  17(4):40&ndash;52.
- Tsantalis, N., Chatzigeorgiou, A., Stephanides, G., & Halkidis, S. T.
  (2006). Design pattern detection using similarity scoring.
  *IEEE Transactions on Software Engineering* 32(11):896&ndash;909.
  doi:10.1109/TSE.2006.112 (pattern4 detector).
- Arcelli Fontana, F., Pigazzini, I., Roveda, R., Tamburri, D.,
  Zanoni, M., & Lenarduzzi, V. (2017). Arcan: A tool for
  architectural smells detection. *ECSA 2017* — documented fallback
  detector with compatible output schema.
- MacCormack, A., Rusnak, J., & Baldwin, C. Y. (2007). Exploring the
  structure of complex software designs: An empirical study of open
  source and proprietary code. *Management Science* 52(7):1015&ndash;1030.
