---
title: "Lift — Brooksq Model from Apache Helix"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Brooksq Szz Injection Leak}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu)
require(data.table)
require(stringi)
require(knitr)
require(yaml)
require(magrittr)
# Helpers (myths_*.R) loaded with library(kaiaulu).
```

# Introduction

The `brooksq` model extends Brooks with quality tracking: a late-hire
boost slows veteran velocity AND raises the bug injection rate, so
quality-adjusted progress `y = Done - 5*Esc` is hurt twice over. The
sd.py model declares `boost` as its `ctrl` variable; the thesis says
y drops when boost increases.

This notebook lifts the quality side of `brooksq` from Apache Helix
using B-SZZ output (PyDriller). The Brooks side (`brooks_tax`) is
reused from `lift_brooks.Rmd`.

Calibration outputs:

| sd.py param   | source                                           |
|---------------|--------------------------------------------------|
| `brooks_tax`  | veteran velocity drop in post-hire window        |
| `inj_rate`    | bug-introducing commits per day, pre-hire window |
| `inj_rate'`   | same, post-hire — the boost-driven increase      |
| `leak_rate`   | fraction of bugs unfixed past a latency threshold|

# Project Configuration

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/helix.yml")

perceval_path <- get_tool_project("perceval", tool)
git_repo_path <- get_git_repo_path(conf)

file_extensions    <- get_file_extensions(conf)
substring_filepath <- get_substring_filepath(conf)
```

# Parse Git Log

```{r}
project_git <- parse_gitlog(perceval_path, git_repo_path)

project_git <- project_git %>%
  filter_by_file_extension(file_extensions, "file_pathname") %>%
  filter_by_filepath_substring(substring_filepath, "file_pathname")

project_git$author_datetimetz <- as.POSIXct(project_git$author_datetimetz,
                                            format = "%a %b %d %H:%M:%S %Y %z",
                                            tz = "UTC")
```

# Identity Resolution

```{r}
project_log <- list(git_log = project_git)
name_column <- c("author_name_email")
project_log <- identity_match(project_log,
                              name_column,
                              assign_identity_function = assign_exact_identity,
                              use_name_only = FALSE,
                              label = "identity_id")
project_git <- project_log$git_log
```

# Parse SZZ Bug-Fix Pairs

The SZZ pairs are produced by an external PyDriller pass; the script
lives at `../scripts/szz_helix.py` and the CSV at
`../data/helix/derived/szz_pairs.csv`. Re-run that script whenever the
git repo is refreshed.

```{r}
szz <- parse_szz_bugfixes("../data/helix/derived/szz_pairs.csv")

nrow(szz)
length(unique(szz$fixing_commit_hash))
length(unique(szz$introducing_commit_hash))
```

# Detect Late-Hire Events

```{r}
late_hires <- detect_late_hires(project_git,
                                min_project_age_days = 365)
kable(head(late_hires, 20))
```

# Brooks Side — Veteran Velocity Drop

```{r}
window_days <- 90

velocity_changes <- compute_velocity_changes(project_git,
                                             late_hires,
                                             window_days = window_days)
velocity_changes[, brooks_tax := (pre_velocity - post_velocity) / pre_velocity]

summary(velocity_changes$brooks_tax)
```

# Quality Side — Bug Injection Around Each Hire

For each late hire, count bug-introducing commits in the pre- and
post-hire windows. A positive `inj_rate_increase` confirms brooksq's
quality thesis: hires don't just slow veterans, they introduce bugs.

```{r}
injection_changes <- compute_injection_changes(szz,
                                               late_hires,
                                               window_days = window_days)
injection_changes[, inj_rate_increase := inj_rate_post - inj_rate_pre]

kable(head(injection_changes, 20))
summary(injection_changes$inj_rate_increase)
```

# Leak Rate

Fraction of bug-introducing commits whose fix takes longer than
`latency_days`. Higher = more bugs leak past the immediate-review
window into `Esc`.

```{r}
latency_days <- 30
leak_rate    <- estimate_leak_rate(szz, latency_days = latency_days)
leak_rate
```

# Sanity Checks

Two-part sanity check:

**(1) Bug-count dependency**: this lift DOES depend on bug attribution.
We use the JIRA-key-in-commit-message heuristic to identify bug-fix
commits (see `scripts/szz_helix.py`). Datasets without JIRA keys in
commit messages — pure-GitHub-issue projects — would need a different
seed mechanism (e.g. linked PR closes).

**(2) Identity bridging**: the brooksq lift uses identity_match on
git only. SZZ pairs are joined on `commit_hash`, not on developer.
GitHub would still need an extra alias source if we extended this
to author-attributed injection rates.

# Outputs for the SD Framework

```{r}
out <- data.table(
  project           = "helix",
  brooks_tax_median = median(velocity_changes$brooks_tax,        na.rm = TRUE),
  inj_rate_pre_med  = median(injection_changes$inj_rate_pre,     na.rm = TRUE),
  inj_rate_post_med = median(injection_changes$inj_rate_post,    na.rm = TRUE),
  inj_rate_increase = median(injection_changes$inj_rate_increase,na.rm = TRUE),
  leak_rate         = leak_rate,
  n_hires           = nrow(velocity_changes),
  n_szz_pairs       = nrow(szz),
  window_days       = window_days,
  latency_days      = latency_days,
  seed              = seed
)
fwrite(out, "../outputs/lift_brooksq_helix.csv")
kable(out)
```

Verdict interpretation: brooksq's thesis is supported if both
`brooks_tax_median > 0` AND `inj_rate_increase > 0`. Falsified if
either is at-or-below zero. Pass these calibrated values to
`sd.opt(brooksq, ...)` and re-run `verdict()` to obtain the y0 vs y1
margin on the calibrated model.
