---
title: "Lift — Brooks Model from Apache Helix"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Brooks Late Hire Velocity}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu)
require(data.table)
require(stringi)
require(knitr)
require(yaml)
require(magrittr)
# Helpers (myths_*.R) loaded with library(kaiaulu).
```

# Introduction

The Brooks model expresses the central thesis of *The Mythical
Man-Month* (Brooks 1975): adding people to a late project makes it
later. In SD form, two stocks track personnel — veteran developers
(`Vet`) and new hires (`New`) — and a productivity tax falls on
veterans whenever newcomers join (training cost + intra-team
communication overhead). The SD model is in the downstream SD model.

This notebook lifts the model's productivity-tax parameter from
Apache Helix's git history. We need only the git log and kaiaulu's
`identity_match` (no proprietary tooling).

**Lift output**: a calibrated value of `brooks_tax` =
`(pre_velocity - post_velocity) / pre_velocity`, where the velocities
are veteran commit rates in symmetric windows before and after each
late-hire event. A positive median supports Brooks; a negative or
near-zero median is itself a useful falsification — late hires
*accelerated* veterans on this project.

**Method outline**:
1. Parse git log via `parse_gitlog(perceval, repo)`.
2. Resolve developer aliases via `identity_match(...)`.
3. Detect late-hire events (first-commit ≥ 365 days after project
   start) with `detect_late_hires()`.
4. For each late hire, compute pre/post veteran commit rates with
   `compute_velocity_changes()`.
5. Aggregate `brooks_tax` median + mean.

The two helper functions are in `lifts/functions.R`; both are
candidates for upstream contribution into `kaiaulu/R/`.

# Project Configuration

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/helix.yml")

perceval_path <- get_tool_project("perceval", tool)
git_repo_path <- get_git_repo_path(conf)

file_extensions      <- get_file_extensions(conf)
substring_filepath   <- get_substring_filepath(conf)
```

# Parse Git Log

```{r}
project_git <- parse_gitlog(perceval_path, git_repo_path)

project_git <- project_git %>%
  filter_by_file_extension(file_extensions, "file_pathname") %>%
  filter_by_filepath_substring(substring_filepath, "file_pathname")

project_git$author_datetimetz <- as.POSIXct(project_git$author_datetimetz,
                                            format = "%a %b %d %H:%M:%S %Y %z",
                                            tz = "UTC")
```

Sanity check the basic shape:

```{r}
nrow(project_git)
length(unique(project_git$commit_hash))
length(unique(project_git$author_name_email))
```

# Identity Resolution

Merges `(name, email)` aliases so one human contributor maps to one
developer record. Standard kaiaulu step via
`\code{\link{identity_match}}` even when working with a single
source, since downstream cohort logic operates on `identity_id`.

```{r}
project_log <- list(git_log = project_git)
name_column <- c("author_name_email")
project_log <- identity_match(project_log,
                              name_column,
                              assign_identity_function = assign_exact_identity,
                              use_name_only = FALSE,
                              label = "identity_id")
project_git <- project_log$git_log
```

After identity_match, each row has `identity_id` and `raw_name`. Use
`identity_id` for all downstream cohort analysis.

# Detect Late-Hire Events

A late hire is a developer whose first commit is at least
`min_project_age_days` after project start. Function defined in
`functions.R`:

```{r}
project_start <- min(project_git$author_datetimetz)
late_hires    <- detect_late_hires(project_git,
                                   min_project_age_days = 365)
kable(head(late_hires, 20))
```

# Compute Veteran Velocity Before and After Each Hire

Window size in days:

```{r}
window_days <- 90
```

```{r}
velocity_changes <- compute_velocity_changes(project_git,
                                             late_hires,
                                             window_days = window_days)
kable(velocity_changes)
```

# Compute brooks_tax

```{r}
brooks_tax_values <- velocity_changes[,
  brooks_tax := (pre_velocity - post_velocity) / pre_velocity
]

summary(brooks_tax_values$brooks_tax)
```

Median brooks_tax is the calibration value for the downstream SD model.
A positive median supports Brooks; a negative or near-zero median is
itself an interesting falsification.

# Verdict on Helix

The median `brooks_tax` (final cell, below) is **positive on Helix**
(≈ 0.11). Veterans lose roughly 11% of their commit velocity in the
90 days after a new committer joins. The size is modest but the
direction matches Brooks's claim. Family-member tests on 7 more
projects (in `findings.md` F3) show the same sign for 5 of them, with
magnitudes 0.03 → 0.31 — large project-to-project variance suggests
team-size and mentoring practices modulate the effect heavily.

# References

- Brooks, F. P. (1975). *The Mythical Man-Month*. Addison-Wesley.
- Curtis, B., Krasner, H., & Iscoe, N. (1988). A field study of the
  software design process for large systems. *Communications of the ACM*
  31(11):1268&ndash;1287. doi:10.1145/50087.50089.
- Replication: re-run this notebook on a different Apache project by
  swapping `../conf/helix.yml` for `../conf/<project>.yml`.

# Sanity Checks

**Bug-count dependency**: this lift does not use bug count; the
choice of communication source (mbox / github / jira) is irrelevant.
The recipe runs on any git-only project.

**Identity bridging**: `\code{\link{identity_match}}` is applied on
a single source (git). No comms-to-code bridging required. If the
brooks recipe were extended to discriminate bug-introducing
late-hires from feature-introducing late-hires, JIRA + git
identity_match would be required, and a GitHub username alias
source would also be needed.

# Outputs for the SD Framework

Write the calibrated `brooks_tax` value(s) to a CSV the Python
framework can consume:

```{r}
out <- data.table(
  project          = "helix",
  brooks_tax_median = median(brooks_tax_values$brooks_tax, na.rm = TRUE),
  brooks_tax_mean   = mean(brooks_tax_values$brooks_tax,   na.rm = TRUE),
  n_hires          = nrow(velocity_changes),
  window_days      = window_days,
  seed             = seed
)
fwrite(out, "../outputs/lift_brooks_helix.csv")
kable(out)
```
