---
title: "Lift -- Brooks Tax from Pre-Extracted Commits CSV"
output:
  html_document:
    toc: true
    number_sections: true
params:
  commits: "commits.csv"   # input: extract output (kaiaulu::parse_gitlog)
  out:     "brooks.csv"    # output: one-row lift CSV (brooks_tax median + counts)
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Brooks Lift From Commits}
  %\VignetteEncoding{UTF-8}
---

```{r}
set.seed(1)
```

```{r warning=FALSE, message=FALSE}
require(data.table)
require(stringi)
```

# Method

Reads a pre-extracted commits CSV (output of `kaiaulu::parse_gitlog`)
and computes the Brooks tax = (pre_velocity - post_velocity) /
pre_velocity, where velocities are veteran commit rates in symmetric
90-day windows around each late-hire event.

**Late hire** = an author whose first commit lands more than 365 days
after the project's first commit.

# Lift

```{r}
commits <- fread(params$commits)
setnames(commits, tolower(names(commits)))

# kaiaulu schema: author_name_email, author_datetimetz, commit_hash, ...
commits[, date := as.POSIXct(author_datetimetz,
                             format = "%a %b %d %H:%M:%S %Y %z",
                             tz = "UTC")]
if (all(is.na(commits$date))) {
  commits[, date := as.POSIXct(author_datetimetz, tz = "UTC")]
}
commits <- commits[!is.na(date)]
setnames(commits, "author_name_email", "author")
setorder(commits, date)

project_start <- min(commits$date)
WINDOW_DAYS   <- 90
LATE_THRESH   <- 365

# Per-author first commit (dedupe across multi-file commit rows).
first_by <- unique(commits, by = c("commit_hash", "author"))[, .(first = min(date)), by = author]
first_by[, days_into := as.numeric(difftime(first, project_start, units = "days"))]
late_hires <- first_by[days_into >= LATE_THRESH]

# For each late-hire event, count VETERAN commits in [event-90d, event)
# vs [event, event+90d).  A veteran = author whose first commit is
# before this event.
tax_rows <- list()
# Count VETERAN COMMITS (not commit-file rows); dedupe by hash.
veteran_commits <- unique(commits, by = "commit_hash")
for (i in seq_len(nrow(late_hires))) {
  ev <- late_hires$first[i]
  vets <- first_by[first < ev, author]
  pre  <- nrow(veteran_commits[author %in% vets & date >= ev - WINDOW_DAYS * 86400 & date <  ev])
  post <- nrow(veteran_commits[author %in% vets & date >= ev                  & date <  ev + WINDOW_DAYS * 86400])
  if (pre > 0) {
    tax_rows[[i]] <- data.table(event = ev, pre = pre, post = post,
                                tax = (pre - post) / pre)
  }
}
taxes <- rbindlist(tax_rows)

out <- data.table(
  metric        = c("brooks_tax_median", "brooks_tax_mean"),
  value         = c(median(taxes$tax, na.rm = TRUE), mean(taxes$tax, na.rm = TRUE)),
  n_late_hires  = nrow(late_hires),
  n_paired      = nrow(taxes),
  n_commits     = nrow(commits))
fwrite(out, params$out)
out
```

# Notes

* `brooks_tax > 0` supports Brooks (veterans slowed after the hire).
* `brooks_tax < 0` falsifies it for this project (hire correlated with
  acceleration).
* The 365-day late-hire threshold + 90-day window are kaiaulu's
  defaults; both can be tightened by editing this vignette.
