---
title: "Lift — Bugs Model (Goel-Okumoto) from Apache Helix"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Bugs Gokumoto Growth}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu); require(data.table); require(stringi)
require(knitr); require(yaml); require(jsonlite)
# Helpers (myths_*.R) loaded with library(kaiaulu).
```

# Introduction

The `bugs` model (Goel & Okumoto 1979) is the canonical software-
reliability-growth thesis: cumulative bug detection follows an
exponential approach to an asymptotic total. The SD form is in
`models/sd.py:bugs()`:

$$N(t) = a \cdot (1 - e^{-b \cdot t})$$

where `a` is the eventual total bug population and `b` is the
discovery rate.

**Input source choice**: bugs needs a bug-count signal with
timestamps. Helix provides three candidates:

- **GitHub issue dump** (label='bug', closed_at) — used here
- JIRA dump (issuetype='Bug', resolutiondate) — partial dump only;
  the available window has 0 Bug-type issues
- commit-message heuristic — too noisy without ground-truth filter

We choose GitHub issues because the dump is complete (1879 issues)
and the `bug` label provides a clean classifier.

**Method outline**:
1. Parse all GitHub issue JSON files.
2. Filter issues with label='bug' AND non-null `closed_at`.
3. Sort by close time. Build cumulative-N(t) trajectory.
4. Grid-search Goel-Okumoto fit (a × b parameter pairs).
5. Output: `a`, `b`, R² fit quality.

# Configuration

```{r}
gh_issue_glob <- "/Users/timm/Downloads/helix/github/apache_helix/issue/*.json"
```

# Load GitHub Issues

```{r}
files <- Sys.glob(gh_issue_glob)
cat("Issue files:", length(files), "\n")

issues <- rbindlist(lapply(files, function(f) {
  d <- fromJSON(f, simplifyVector = FALSE)
  if (is.list(d) && !is.null(d[[1]]$id)) {
    # list of issues
    rbindlist(lapply(d, function(i) {
      labels <- vapply(i$labels %||% list(), `[[`, "", "name")
      data.table(
        id        = as.integer(i$id %||% NA),
        state     = i$state %||% NA_character_,
        labels    = paste(labels, collapse = ";"),
        closed_at = i$closed_at %||% NA_character_
      )
    }), fill = TRUE)
  } else if (is.list(d) && "issues" %in% names(d)) {
    rbindlist(lapply(d$issues, function(i) data.table(
      id = as.integer(i$id %||% NA),
      state = i$state %||% NA_character_,
      labels = "",
      closed_at = i$closed_at %||% NA_character_
    )), fill = TRUE)
  } else NULL
}), fill = TRUE)

cat("Total issues parsed:", nrow(issues), "\n")
cat("States:\n"); print(table(issues$state))
```

# Filter to closed bug-labeled issues

```{r}
bugs <- issues[grepl("bug", labels, ignore.case = TRUE) &
               !is.na(closed_at) & closed_at != ""]
cat("Bug-labeled + closed:", nrow(bugs), "\n")
bugs[, closed_ts := as.POSIXct(closed_at, format = "%Y-%m-%dT%H:%M:%SZ",
                               tz = "UTC")]
setorder(bugs, closed_ts)
```

# Fit Goel-Okumoto

```{r}
t0       <- min(bugs$closed_ts)
rel_t    <- as.numeric(difftime(bugs$closed_ts, t0, units = "secs"))
cum_N    <- seq_along(rel_t)
mean_N   <- mean(cum_N)
ss_tot   <- sum((cum_N - mean_N)^2)
a_max    <- max(cum_N) * 1.2
best     <- list(a = NA, b = NA, sse = Inf)
for (a in a_max * c(0.8, 1.0, 1.2, 1.5, 2.0, 3.0)) {
  for (b in c(1e-9, 1e-8, 1e-7, 1e-6, 1e-5, 1e-4)) {
    pred <- a * (1 - exp(-b * rel_t))
    sse  <- sum((pred - cum_N)^2)
    if (sse < best$sse) best <- list(a = a, b = b, sse = sse)
  }
}
r2 <- 1 - best$sse / ss_tot
cat(sprintf("Goel-Okumoto: a=%.2f  b=%.3e  R^2=%.4f\n",
            best$a, best$b, r2))
span_days <- as.numeric(difftime(max(bugs$closed_ts), t0, units = "days"))
cat(sprintf("Span: %.1f days\n", span_days))
```

# Outputs

```{r}
dir.create("../outputs", showWarnings = FALSE)
out <- data.table(
  project         = "helix",
  n_issues_total  = nrow(issues),
  n_bugs_closed   = nrow(bugs),
  gokumoto_a      = best$a,
  gokumoto_b      = best$b,
  fit_r2          = r2,
  span_days       = span_days,
  seed            = seed
)
fwrite(out, "../outputs/lift_bugs_helix.csv")
kable(out)
```

# Verdict on Helix

The Goel-Okumoto fit predicts an eventual total of ~245 closed bugs
(`a`) on a discovery rate of `b ≈ 1e-8 per second`. The R² of ~0.62
indicates a moderate fit — the trajectory deviates from pure
exponential, suggesting Helix is still in the early discovery regime
where new sub-systems keep entering the bug-detectable population.

In `models/sd.py:bugs()` terms: the initial `Latent` stock would be
≈ 245 (the asymptote), of which 170 have been `Found` to date and
all 170 are `Fixed`. The model's RQ is "2x initial Latent → ~2x
eventual Fixed (linearity)" — Helix is the wrong test bed because
the trajectory hasn't saturated yet. A family-member test on a
project at or near saturation (e.g. an older project with declining
bug arrivals) would be a better falsification target.

# Family-Member Test

Three projects with bug data on disk:

| project  | n_bugs | source     | a (asymptote) | b           | R²   |
|----------|-------:|------------|--------------:|-------------|-----:|
| helix    | 170    | GH labels  | 244.80        | 1.00e-08    | 0.62 |
| kaiaulu  | 22     | GH labels  | 26.40         | 1.00e-08    | 0.91 |
| camel    | 185    | JIRA       | 177.60        | 1.00e-07    | 0.60 |

kaiaulu's higher R² is partly small-sample artifact (fewer outliers).
camel's `b` is 10x kaiaulu's — faster discovery in the older / busier
project. Other 5 projects don't have a bug-classification source on
disk yet.

# References

- Goel, A. L. & Okumoto, K. (1979). Time-dependent error-detection
  rate model for software reliability and other performance measures.
  *IEEE Transactions on Reliability*, R-28(3), 206–211.
- `models/sd.py:bugs()` — the SD model.
- `scripts/lift_bugs_gh.py` + `scripts/lift_bugs_jira.py` — Python
  variants for the multi-project sweep that produced the family-
  member table above.

# Notes

The Goel-Okumoto fit is **grid-search**, not LM-style optimisation.
Six values of `a`, six of `b`. Adequate for a sanity calibration, not
publication-grade. For paper use, replace with `nls()` or `optim()`
with bounds.
