---
title: "Custom Graph Weight Showcase"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Custom Graph Weight Showcase}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE,message=FALSE}
require(kaiaulu)
require(igraph)
require(visNetwork)
require(data.table)
require(yaml)
require(stringi)
require(knitr)
```

# Introduction

This notebook demonstrates how to assign custom numeric values as edge weights when transforming raw data into a graph object. Specifically, it uses sentiment polarity scores from GitHub issue comments as edge weights in a bipartite author-to-issue network, and visualizes the resulting network with edges colored by sentiment.

Polarity integers (`0` = neutral, `1` = positive, `2` = negative) are remapped to (`0`, `1`, `-1`) before being passed to `transform_reply_to_bipartite_network()` via the `weight` parameter. The `weight_agg` parameter controls how multiple comments between the same author and issue are aggregated — defaulting to `mean`.

# Load Data

The input is a CSV of Kaiaulu GitHub issue comments with predicted sentiment polarity labels.

```{r}
dt <- fread("https://github.com/user-attachments/files/27417102/kaiaulu_predictions.csv")
kable(head(dt[, .(comment_id, issue_url, comment_user_login, polarity)]))
```

# Prepare Data

Rename columns to match the expected `reply_from` and `reply_subject` format, and remap polarity integers (0 = neutral, 1 = positive, 2 = negative) to signed weights (0, 1, -1).

```{r}
dt[, reply_from    := comment_user_login]
dt[, reply_subject := sub(".*/issues/", "#", issue_url)]
dt[, polarity_int  := fcase(
  polarity == 1L,  1L,
  polarity == 0L,  0L,
  polarity == 2L, -1L
)]
kable(head(dt[, .(reply_from, reply_subject, polarity_int)]))
```

# Build Weighted Bipartite Network

Transform the reply table into a bipartite graph using sentiment polarity as the edge weight, aggregated by mean across multiple comments between the same author and issue.

```{r}
reply_graph <- transform_reply_to_bipartite_network(dt, weight = "polarity_int", weight_agg = mean)
kable(head(reply_graph[["nodes"]]))
kable(head(reply_graph[["edgelist"]]))
```

# Visualize Network

Edge color reflects the mean sentiment between an author and a commit: green for positive, black for neutral, and red for negative.

```{r}
nodes <- reply_graph[["nodes"]]
edges <- reply_graph[["edgelist"]]

edges[, color := fcase(
  weight > 0,  "green",
  weight == 0, "black",
  weight < 0,  "red"
)]

is_directed <- any(edges[["direction"]] == "directed")
plot_reply_graph <- igraph::graph_from_data_frame(d = edges,
                                                  directed = is_directed,
                                                  vertices = nodes)
visIgraph(plot_reply_graph, randomSeed = 1)
```

# Emotion Graph

This section demonstrates using binary emotion labels as edge weights in a bipartite author-to-issue network. The dataset contains JIRA comments from the Apache Harmony project, each annotated with six binary emotion labels: `love`, `joy`, `sadness`, `anger`, `surprise`, and `fear`.

Each emotion is visualized as a separate graph. The `weight` parameter is set to the emotion column, and `weight_agg = mean` aggregates multiple comments between the same author and issue into a proportion (0 = no comments with that emotion, 1 = all comments with that emotion). Only edges with weight > 0 are shown. Edge color encodes intensity using three levels — light, medium, dark — and edge labels show the rounded weight value.

## Load Emotion Data

```{r}
emotion_dt <- as.data.table(read.csv("https://github.com/user-attachments/files/27313854/harmony_emotion_comments_joined.csv"))
kable(head(emotion_dt[, .(comment_id, issue_key, author_login, author_name, love, joy, sadness, anger, surprise, fear)]))
```

## Prepare Data

```{r}
emotion_dt[, reply_from    := author_login]
emotion_dt[, reply_subject := issue_key]
kable(head(emotion_dt[, .(reply_from, reply_subject, love, joy, sadness, anger, surprise, fear)]))
```

## Love

```{r fig.width=10, fig.height=7}
love_graph <- transform_reply_to_bipartite_network(emotion_dt, weight = "love", weight_agg = mean)
nodes <- love_graph[["nodes"]]
edges <- love_graph[["edgelist"]]
if (nrow(edges) > 0) {
  edges[, color := "hotpink"]
  edges[, label := as.character(round(weight, 2))]
  is_directed <- any(love_graph[["edgelist"]][["direction"]] == "directed")
  visIgraph(igraph::graph_from_data_frame(d = edges, directed = is_directed, vertices = nodes), randomSeed = seed) |>
    visEdges(font = list(align = "horizontal", background = "white", strokeWidth = 0))
}
```

## Joy

```{r fig.width=10, fig.height=7}
joy_graph <- transform_reply_to_bipartite_network(emotion_dt, weight = "joy", weight_agg = mean)
nodes <- joy_graph[["nodes"]]
edges <- joy_graph[["edgelist"]]
if (nrow(edges) > 0) {
  edges[, color := "darkorange"]
  edges[, label := as.character(round(weight, 2))]
  is_directed <- any(joy_graph[["edgelist"]][["direction"]] == "directed")
  visIgraph(igraph::graph_from_data_frame(d = edges, directed = is_directed, vertices = nodes), randomSeed = seed) |>
    visEdges(font = list(align = "horizontal", background = "white", strokeWidth = 0))
}
```

## Sadness

```{r fig.width=10, fig.height=7}
sadness_graph <- transform_reply_to_bipartite_network(emotion_dt, weight = "sadness", weight_agg = mean)
nodes <- sadness_graph[["nodes"]]
edges <- sadness_graph[["edgelist"]]
if (nrow(edges) > 0) {
  edges[, color := "steelblue"]
  edges[, label := as.character(round(weight, 2))]
  is_directed <- any(sadness_graph[["edgelist"]][["direction"]] == "directed")
  visIgraph(igraph::graph_from_data_frame(d = edges, directed = is_directed, vertices = nodes), randomSeed = seed) |>
    visEdges(font = list(align = "horizontal", background = "white", strokeWidth = 0))
}
```

## Anger

```{r fig.width=10, fig.height=7}
anger_graph <- transform_reply_to_bipartite_network(emotion_dt, weight = "anger", weight_agg = mean)
nodes <- anger_graph[["nodes"]]
edges <- anger_graph[["edgelist"]]
if (nrow(edges) > 0) {
  edges[, color := "darkred"]
  edges[, label := as.character(round(weight, 2))]
  is_directed <- any(anger_graph[["edgelist"]][["direction"]] == "directed")
  visIgraph(igraph::graph_from_data_frame(d = edges, directed = is_directed, vertices = nodes), randomSeed = seed) |>
    visEdges(font = list(align = "horizontal", background = "white", strokeWidth = 0))
}
```

## Surprise

```{r fig.width=10, fig.height=7}
surprise_graph <- transform_reply_to_bipartite_network(emotion_dt, weight = "surprise", weight_agg = mean)
nodes <- surprise_graph[["nodes"]]
edges <- surprise_graph[["edgelist"]]
if (nrow(edges) > 0) {
  edges[, color := "darkcyan"]
  edges[, label := as.character(round(weight, 2))]
  is_directed <- any(surprise_graph[["edgelist"]][["direction"]] == "directed")
  visIgraph(igraph::graph_from_data_frame(d = edges, directed = is_directed, vertices = nodes), randomSeed = seed) |>
    visEdges(font = list(align = "horizontal", background = "white", strokeWidth = 0))
}
```

## Fear

```{r fig.width=10, fig.height=7}
fear_graph <- transform_reply_to_bipartite_network(emotion_dt, weight = "fear", weight_agg = mean)
nodes <- fear_graph[["nodes"]]
edges <- fear_graph[["edgelist"]]
if (nrow(edges) > 0) {
  edges[, color := "mediumpurple"]
  edges[, label := as.character(round(weight, 2))]
  is_directed <- any(fear_graph[["edgelist"]][["direction"]] == "directed")
  visIgraph(igraph::graph_from_data_frame(d = edges, directed = is_directed, vertices = nodes), randomSeed = seed) |>
    visEdges(font = list(align = "horizontal", background = "white", strokeWidth = 0))
}
```
