---
title: "Custom Graph Weight Showcase"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Custom Graph Weight Showcase}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE,message=FALSE}
require(kaiaulu)
require(igraph)
require(visNetwork)
require(data.table)
require(yaml)
require(stringi)
require(knitr)
```

# Introduction

This notebook demonstrates how to assign custom numeric values as edge weights when transforming raw data into a graph object. Specifically, it uses sentiment polarity scores from GitHub commit comments as edge weights in a bipartite author-to-commit network, and visualizes the resulting network with edges colored by sentiment.

Sentiment labels (`positive`, `neutral`, `negative`) are first converted to integers (`1`, `0`, `-1`) before being passed to `transform_reply_to_bipartite_network()` via the `weight` parameter. The `weight_agg` parameter controls how multiple comments between the same author and commit are aggregated — defaulting to `mean`.

# Load Data

The input is a CSV of CakePHP GitHub commit comments joined with sentiment labels from the MSR14 gold standard dataset.

```{r}
dt <- fread("rawdata/github/cakephp_sentiment_commit_comments_joined.csv")
kable(head(dt))
```

# Prepare Data

Rename columns to match the expected `reply_from` and `reply_subject` format, and convert polarity strings to integers.

```{r}
dt[, reply_from    := author_login_kaiaulu]
dt[, reply_subject := commit_id]
dt[, polarity_int  := fcase(
  polarity == "positive",  1L,
  polarity == "neutral",   0L,
  polarity == "negative", -1L
)]
kable(head(dt[, .(reply_from, reply_subject, polarity_int)]))
```

# Build Weighted Bipartite Network

Transform the reply table into a bipartite graph using sentiment polarity as the edge weight, aggregated by mean across multiple comments between the same author and commit.

```{r}
reply_graph <- transform_reply_to_bipartite_network(dt, weight = "polarity_int", weight_agg = mean)
kable(head(reply_graph[["nodes"]]))
kable(head(reply_graph[["edgelist"]]))
```

# Visualize Network

Edge color reflects the mean sentiment between an author and a commit: green for positive, black for neutral, and red for negative.

```{r}
nodes <- reply_graph[["nodes"]]
edges <- reply_graph[["edgelist"]]

edges[, color := fcase(
  weight > 0,  "green",
  weight == 0, "black",
  weight < 0,  "red"
)]

is_directed <- any(edges[["direction"]] == "directed")
plot_reply_graph <- igraph::graph_from_data_frame(d = edges,
                                                  directed = is_directed,
                                                  vertices = nodes)
visIgraph(plot_reply_graph, randomSeed = 1)
```
