---
title: "Lift — Debt Model from Apache Helix"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Debt Refactor Pay Rate}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu)
require(data.table)
require(stringi)
require(knitr)
require(yaml)
require(magrittr)
# Helpers (myths_*.R) loaded with library(kaiaulu).
```

# Introduction

The `debt` model (Cunningham 1992) expresses the thesis that shipping
fast accrues technical debt, which then slows future shipping. The SD
form has stocks `Feat` (cumulative features), `Debt` (current debt),
`Vel` (velocity); parameters `born_rate` (debt born per ship),
`intr_rate` (compounding interest on existing debt), `pay_rate` (debt
paid by refactoring).

This notebook lifts `debt` from Apache Helix using:

| sd.py param  | source                                            |
|--------------|---------------------------------------------------|
| `pay_rate`   | RefactoringMiner: refactor commits / total commits|
| `born_rate`  | gitlog: share of "big" multi-file commits         |
| `intr_rate`  | proxy: bug-touch revisit rate on prior-changed files (TODO) |

`intr_rate` remains imperfect from git-alone; either accept fitting via
`sd.opt()` or refine later using SZZ pairs joined to gitlog churn.

# Project Configuration

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/helix.yml")

perceval_path  <- get_tool_project("perceval", tool)
refminer_path  <- tool$refactoring_miner$executable
git_repo_path  <- get_git_repo_path(conf)

file_extensions    <- get_file_extensions(conf)
substring_filepath <- get_substring_filepath(conf)
```

# Parse Git Log

```{r}
project_git <- parse_gitlog(perceval_path, git_repo_path)

project_git <- project_git %>%
  filter_by_file_extension(file_extensions, "file_pathname") %>%
  filter_by_filepath_substring(substring_filepath, "file_pathname")

project_git$author_datetimetz <- as.POSIXct(project_git$author_datetimetz,
                                            format = "%a %b %d %H:%M:%S %Y %z",
                                            tz = "UTC")
```

# Load RefactoringMiner Output

The full RefactoringMiner pass on Helix produces JSON at
`../data/helix/derived/refminer.json`. Re-run via:

```bash
tools/RefactoringMiner-3.0.10/bin/RefactoringMiner -a \
  data/helix/git_repo master \
  -json data/helix/derived/refminer.json
```

```{r}
refminer_json <- "../data/helix/derived/refminer.json"
refactorings  <- flatten_refactoring_json(refminer_json)

nrow(refactorings)
length(unique(refactorings$commit_hash))
table(refactorings$refactoring_type)[1:10]
```

# Compute pay_rate per Window

```{r}
window_days <- 90
pay_rate_table <- compute_pay_rate(project_git, refactorings,
                                   window_days = window_days)
kable(head(pay_rate_table, 20))
summary(pay_rate_table$pay_rate)
```

# Compute born_rate Proxy per Window

```{r}
born_rate_table <- compute_born_rate_proxy(project_git,
                                           window_days = window_days,
                                           big_commit_files = 5)
kable(head(born_rate_table, 20))
summary(born_rate_table$born_rate)
```

# Sanity Checks

**(1) Bug-count dependency**: this lift does NOT depend on bug-count,
so JIRA/GitHub/mbox choice is irrelevant. The lift would work on any
Java-or-similar project with a git history.

**(2) Identity bridging**: `debt` is whole-project, not per-developer.
`identity_match` is not required for the lift, though it would be for
any extension that breaks pay_rate by developer cohort.

# Outputs for the SD Framework

```{r}
out <- data.table(
  project           = "helix",
  pay_rate_median   = median(pay_rate_table$pay_rate,   na.rm = TRUE),
  pay_rate_mean     = mean(pay_rate_table$pay_rate,     na.rm = TRUE),
  born_rate_median  = median(born_rate_table$born_rate, na.rm = TRUE),
  born_rate_mean    = mean(born_rate_table$born_rate,   na.rm = TRUE),
  n_refactor_events = nrow(refactorings),
  n_commits         = uniqueN(project_git$commit_hash),
  window_days       = window_days,
  seed              = seed
)
fwrite(out, "../outputs/lift_debt_helix.csv")
kable(out)
```

Verdict interpretation: with `pay_rate` and `born_rate` calibrated,
pass them to `sd.opt(debt, ...)` for `intr_rate` fitting (only net Debt
change is observable from gitlog alone), then re-run `verdict()` to
compare baseline (Debt=0) vs perturbed (Debt=50) trajectories.
