---
title: "Lift — Defmap Model from Apache Helix"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Defmap Active Latent}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu); require(data.table); require(stringi)
require(knitr); require(yaml); require(magrittr)
# Helpers (myths_*.R) loaded with library(kaiaulu).
```

# Introduction

The `defmap` model (Abdel-Hamid & Madnick 1991 defect submodel) tracks
Injected / Caught / Latent / Prod stocks. `ctrl` is `tst` (testing
intensity). Thesis: `tst` 2.5 → 0.5 increases operational defects.

This notebook lifts the model's defect-flow stocks across Helix
release-tag phases. Each phase = (tag_i, tag_{i+1}].

# Project Configuration

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/helix.yml")
git_repo_path <- get_git_repo_path(conf)
perceval_path <- get_tool_project("perceval", tool)
```

# Parse Git Log + SZZ

```{r}
project_git <- parse_gitlog(perceval_path, git_repo_path) %>%
  filter_by_file_extension(get_file_extensions(conf), "file_pathname") %>%
  filter_by_filepath_substring(get_substring_filepath(conf), "file_pathname")
project_git$author_datetimetz <- as.POSIXct(project_git$author_datetimetz,
                                            format = "%a %b %d %H:%M:%S %Y %z",
                                            tz = "UTC")
szz <- parse_szz_bugfixes("../data/helix/derived/szz_pairs.csv")
```

# Release-Tag Phase Boundaries

```{r}
tag_dates <- get_tag_dates(git_repo_path)
kable(head(tag_dates, 20))
nrow(tag_dates)
```

# Per-Phase Defect Flow

```{r}
phases <- compute_per_phase_defects(szz, project_git, tag_dates)
kable(phases)
summary(phases$tst_proxy)
```

# Outputs

```{r}
dir.create("../outputs", showWarnings = FALSE)
out <- data.table(project       = "helix",
                  tst_median    = median(phases$tst_proxy, na.rm = TRUE),
                  tst_mean      = mean(phases$tst_proxy,   na.rm = TRUE),
                  total_inject  = sum(phases$injected,  na.rm = TRUE),
                  total_caught  = sum(phases$caught,    na.rm = TRUE),
                  total_leaked  = sum(phases$leaked,    na.rm = TRUE),
                  n_phases      = nrow(phases),
                  seed          = seed)
fwrite(out, "../outputs/lift_defmap_helix.csv")
kable(out)
```
