---
title: "File Networks"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Collaboration Network Parser}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE,message=FALSE}
require(kaiaulu)
require(visNetwork)
require(data.table)
require(stringi)
require(igraph)
require(gh)
require(yaml)
```

```{r}
conf <- yaml::read_yaml("../conf/apr.yml")
depends_jar_path <- conf[["toolpath"]][["depends"]][["bin"]]
language <- conf[["toolpath"]][["depends"]][["codelanguage"]]
git_repo_path <- conf[["datapath"]][["git"]]
```

# Construct File Network

## Project Language

We can use the `gh` package with the language endpoint from Github to figure out what language a project has. Paste the project url below and run the code block.

```{r}
project_github_url <- "https://github.com/apache/apr"


project_github_url <- stri_split_regex(project_github_url,pattern="/")[[1]]
owner <- project_github_url[length(project_github_url)-1]
repo <- project_github_url[length(project_github_url)]
language_distribution_byte <- unlist(gh("GET /repos/:owner/:repo/languages",owner=owner,repo=repo))
language_distribution_byte <- language_distribution_byte/sum(language_distribution_byte)
format(round(language_distribution_byte, 2), nsmall = 2)
```

## Parse Dependencies

Based on the most used language on the project, choose the parameter below for Depends. Accepted languages are: cpp, java, ruby, python, pom.

```{r}
project_dependencies <- parse_dependencies(depends_jar_path,git_repo_path,language=language)
```

A simple filter for `tests` and `__init__.py` is provided in this code block.

```{r}
is_test_src_or_dest <- as.vector(!is.na(stri_match(project_dependencies$src,regex="test")) | 
                                   !is.na(stri_match(project_dependencies$dest,regex="test")))

is_initpy_src_or_dest <- as.vector(!is.na(stri_match_first_regex(project_dependencies$src,
                                                                pattern="__init__.py")) | 
                                   !is.na(stri_match_first_regex(project_dependencies$dest,
                                                                 pattern="__init__.py")))

project_dependencies <- project_dependencies[!is_test_src_or_dest & !is_initpy_src_or_dest]
```

## Example Viz with igraph

```{r}
project_file_network <- parse_dependencies_network(project_dependencies)
project_file_network <- igraph::graph_from_data_frame(d=project_file_network[["edgelist"]], 
                      directed = TRUE, 
                      vertices = project_file_network[["nodes"]])
visIgraph(project_file_network,randomSeed = 1)
```



