---
title: "Lift — DORA Model from Apache Helix"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Dora Szz Metrics}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu); require(data.table); require(stringi)
require(knitr); require(yaml); require(magrittr)
# Helpers (myths_*.R) loaded with library(kaiaulu).
```

# Introduction

The `dora` model (Forsgren, Humble, Kim 2018) tracks Wip / Deploys /
Incidents / Recovery stocks. `ctrl` is `batch_size`. Thesis: large
batches drive change-failure rate (CFR) up, hurting net deploys.

This notebook lifts four DORA metrics from Apache Helix:
- `batch_size`: mean commits between consecutive release tags
- `cfr`: bug-fix commits / total commits
- `arrival_rate`: commits per day
- `rec_rate`: 1 / median bug-fix latency (days)

# Project Configuration

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/helix.yml")
git_repo_path <- get_git_repo_path(conf)
perceval_path <- get_tool_project("perceval", tool)
```

# Parse Git Log + SZZ + Tags

```{r}
project_git <- parse_gitlog(perceval_path, git_repo_path) %>%
  filter_by_file_extension(get_file_extensions(conf), "file_pathname") %>%
  filter_by_filepath_substring(get_substring_filepath(conf), "file_pathname")
project_git$author_datetimetz <- as.POSIXct(project_git$author_datetimetz,
                                            format = "%a %b %d %H:%M:%S %Y %z",
                                            tz = "UTC")
szz       <- parse_szz_bugfixes("../data/helix/derived/szz_pairs.csv")
tag_dates <- get_tag_dates(git_repo_path)
```

# DORA Metrics

```{r}
m <- compute_dora_metrics(szz, project_git, tag_dates)
str(m)
```

# Outputs

```{r}
dir.create("../outputs", showWarnings = FALSE)
out <- data.table(project       = "helix",
                  batch_size    = m$batch_size,
                  cfr           = m$cfr,
                  arrival_rate  = m$arrival_rate,
                  rec_rate      = m$rec_rate,
                  n_tags        = m$n_tags,
                  span_days     = m$span_days,
                  seed          = seed)
fwrite(out, "../outputs/lift_dora_helix.csv")
kable(out)
```

# Verdict Interpretation

`dora` predicts `batch_size = 50` hurts net deploys vs `batch_size =
5`. Helix's measured `batch_size` value here is the operating point;
plug into `sd.opt(dora, ...)` then `verdict()`.
