---
title: "Download Mod Mbox and Pipermail Mailing List Archives"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Download Mod Mbox Mailing List Archives}
  %\VignetteEncoding{UTF-8}
---


```{r}
rm(list = ls())
seed <- 1
set.seed(seed)

# Load libraries
  library(kaiaulu)
  library(data.table)
  library(yaml)
  library(stringi)
  library(XML)
  library(httr)
```


# Introduction

Mailing list data is stored in a variety of archives. See:
- Mod Mbox: [Apache Geronimo](https://geronimo.apache.org/mailing-lists.html)).
- Pipermail: [OpenSSL](https://mta.openssl.org/mailman/listinfo/).
is notebook demonstrates how to download and refresh mailing list archives from Mod Mbox and Pipermail.

## Mailing List Organization

Mailing lists are typically organized by topic or purpose. For example, the [OpenSSL project](https://www.openssl.org/community/mailinglists.html) maintains several mailing lists, each serving a different group:

- **openssl-announce**: For important announcements.
- **openssl-commits**: For commit messages.
- **openssl-project**: For project discussions.
- **openssl-users**: For general user questions and discussions.

Each mailing list maintains archives of past messages, often organized by month and year. These archives can be accessed and downloaded for analysis.

# Project Configuration File 

To start, we load the project configuration file, which contains parameters for downloading the mailing list archives.

// # Project Configuration File

```{r}
conf <- yaml::read_yaml("conf/helix.yml")
mailing_list <- conf[["mailing_list"]][["pipermail"]][["project_key_1"]][["mailing_list"]]
start_year_month <- conf[["mailing_list"]][["pipermail"]][["project_key_1"]][["start_year_month"]]
end_year_month <- conf[["mailing_list"]][["pipermail"]][["project_key_1"]][["end_year_month"]]
save_folder_path <- conf[["mailing_list"]][["pipermail"]][["project_key_1"]][["save_folder_path"]]
```

### Explanation of Configuration Parameters
- mailing_list: The URL of the mailing list archive index page (e.g., https://lists.openssl.org/pipermail/openssl-users/).
- start_year_month: The starting date for downloading archives (in YYYYMM format).
- end_year_month: The ending date for downloading archives (in YYYYMM format).
- save_folder_path: The local directory where the downloaded archives will be saved.


# Pipermail Downloader

```{r}
# Download archives
download_pipermail(
  mailing_list = mailing_list,
  start_year_month = start_year_month,
  end_year_month = end_year_month,
  save_folder_path = save_folder_path
)

```
After running this function, the .mbox files will be saved in the specified directory with filenames like kaiaulu_202310.mbox, kaiaulu_202311.mbox, etc.

