---
title: "Download GitHub Project Issue and Pull Request Comments via API"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Download GitHub Project Issue and Pull Request Comments via API}
  %\VignetteEncoding{UTF-8}
---

# Introduction


In this Vignette, we show how to download GitHub comments from both Issue and Pull Requests. This data may be useful if development communication occurs through issues or pull requests in GitHub in addition or instead of mailing lists. For details on how to merge various communication sources, see the `reply_communication_showcase.Rmd` notebook. The parsed communication table from GitHub comments by `parse_github_comments` has the same format as the mailing list and JIRA comments tables. In turn, the function expects as input the table generated at the end of this notebook.

This notebook assumes you are familiar with obtaining a GitHub token. For details, see the `github_api_showcase.Rmd` notebook. The functions in Kaiaulu will assume you have a token available, which can be passed as parameter. 

```{r warning=FALSE,message=FALSE}
rm(list = ls())
require(kaiaulu)
require(data.table)
require(jsonlite)
require(knitr)
require(magrittr)
require(gt)
```


# GitHub Project's Comments

In this notebook, we are interested in obtaining comment data from the GitHub API. Development communication may occur in either issues or pull requests. [The GitHub API Pulls documentation](https://docs.github.com/en/rest/reference/pulls) states that _'Comments on pull requests can be managed via the Issue Comments API. Every pull request is an issue, but not every issue is a pull request. For this reason, "shared" actions for both features, like manipulating assignees, labels and milestones, are provided within the Issues API.'_

Further details are noted on the issue endpoint: _'Note: GitHub's REST API v3 considers every pull request an issue, but not every issue is a pull request. For this reason, "Issues" endpoints may return both issues and pull requests in the response. You can identify pull requests by the pull_request key. Be aware that the id of a pull request returned from "Issues" endpoints will be an issue id. To find out the pull request id, use the "List pull requests" endpoint.'_

While the above is true for **comments**, the first message of every issue and every pull request (which also include a title) is not regarded by GitHub as a comment. For example, suppose one issue was opened by Author A, and contain only one reply by Author B. In this case, A and B established communication, and we would like this interaction to be reflected in the final table. However, if we were to only use the **comments** endpoint, we would only obtain Author's B comment, and not Author's A. The same is true in Pull Requests.

Therefore, in this Notebook we have to rely on three endpoints from the GitHub API: The `Issue endpoint` to obtain the "first comment" of every issue, the `Pull Request endpoint` to obtain the "first comment" of every pull request, then finally the `Issue and Pull Request Comment endpoint`, which provides comments for both issue and pull requests together.


# Project Configuration File

To use the pipeline, you must specify the organization and project of interest, and your token. Obtain a github token following the instructions [here](https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/creating-a-personal-access-token).

```{r Necessary Parameters}
conf <- parse_config("../conf/kaiaulu.yml")
owner <- get_github_owner(conf, "project_key_1") # Has to match github organization (e.g. github.com/sailuh)
repo <- get_github_repo(conf, "project_key_1") # Has to match github repository (e.g. github.com/sailuh/perceive)

# your file github_token (a text file) contains the GitHub token API
token <- scan("~/.ssh/github_token",what="character",quiet=TRUE)

# The file paths where you want to save the data
save_path_issue_or_pr_comments <- path.expand(get_github_issue_or_pr_comment_path(conf, "project_key_1"))
save_path_issue <- get_github_issue_path(conf, "project_key_1")
save_path_pull_request <- get_github_pull_request_path(conf, "project_key_1")
save_path_commit <- get_github_commit_path(conf, "project_key_1")
save_path_discussions <- get_github_discussions_path(conf, "project_key_1")
```

# Issues

## Issues by Date Range

Kaiaulu offers the ability to download issue data by date range. Using the `github_api_project_issue_by_date()`, you can download issues based on their ['created'](https://docs.github.com/en/search-github/searching-on-github/searching-issues-and-pull-requests#search-by-when-an-issue-or-pull-request-was-created-or-last-updated) dates. The acceptable formats are:

* YYYY-MM-DD
Example: 2023-04-15 (15th April 2023)

* YYYY-MM-DDTHH:MM
Example: 2023-04-15T13:45 (15th April 2023 at 13:45 UTC)


* YYYY-MM-DDTHH:MM:SS
Example: 2023-04-15T13:45:30 (15th April 2023 at 13:45:30 UTC)

* YYYY-MM-DDTHH:MM:SSZ or YYYY-MM-DDTHH:MM:SS+00:00
Example: 2023-04-15T13:45:30Z or 2023-04-15T13:45:30+00:00 (15th April 2023 at 13:45:30 UTC)

If both date range parameters are set to `NULL`, then all issues will be retrieved. Alternatively, if for example `date_lower_bound_issue` is set to "2000-01-01" and `date_upper_bound` is set to "2005-01-01", then only issues created between these two dates will be retrieved.

If you would like to retrieve issues only **after** a certain date, set `date_upper_bound_issue=NULL` and `date_lower_bound` to the date. 

If you would like to retrieve only issues **before** a certain date, set `date_lower_bound_issue=NULL` and date_upper_bound to the date. 


```{r Collect issues by date x, eval = FALSE}
created_lower_bound_issue <- "1990-01-01"
created_upper_bound_issue <- "2021-01-01"
# make initial API CALL
# Acceptable formats for issue_or_pr
# "is:issue"
# "is:pull-request
# issue_or_pr <- "is:issue"
issue_or_pr <- "is:pull-request"
if (issue_or_pr == "is:issue"){
  issue_date_save_path <- save_path_issue_refresh
} else {
  issue_date_save_path <- save_path_pull_request
}
dir.create(issue_date_save_path)
gh_response <- github_api_project_issue_by_date(owner,
                                                repo,
                                                token,
                                                created_lower_bound_issue,
                                                created_upper_bound_issue,
                                                issue_or_pr,
                                                verbose=TRUE)
#dir.create(save_path_issue)

# Make subsequent API calls and write to JSON file along save path
dir.create(save_path_issue_refresh)

github_api_iterate_pages(token,gh_response,
                         issue_date_save_path,
                         prefix="issue",
                         verbose=TRUE)
```

```{r}
# Read all JSON files from the directory
all_issue_files <- lapply(list.files(save_path_issue_refresh, full.names = TRUE), read_json)

# Parse each JSON file using the refresh parser
all_issue <- lapply(all_issue_files, github_parse_search_issues_refresh)


# Combine all the data tables
all_issues_combined <- rbindlist(all_issue, fill = TRUE)

# Display the head of the combined issues table
head(all_issues_combined) %>%
  gt(auto_align = FALSE)
```

## Issues by Customized Queries

You may alternatively have full flexibility on the choice of parameters using `github_api_project_issue()`.

```{r Collect all issues y, eval = FALSE}

query <- "<parameters_of_interest>"

# Acceptable formats for issue_or_pr
# "is:issue"
# "is:pull-request
# issue_or_pr <- "is:issue"
issue_or_pr <- "is:pull-request"
# Check which folder to use
if (issue_or_pr == "is:issue"){
  issue_date_save_path <- save_path_issue_refresh
} else {
  issue_date_save_path <- save_path_pull_request
}

gh_response <- github_api_project_issue_search(owner, repo, token, query, issue_or_pr, verbose=TRUE)

# Make subsequent API calls and write to JSON file along save path

github_api_iterate_pages(token,gh_response,
                         save_path_issue_refresh,
                         prefix="issue",
                         verbose=TRUE)
```

## Refresh Issues

The above function call downloads all the available issues less than the specified limit (default to API limit). The following chunk allows the downloading of issues that have been created only after the most recently created issue already downloaded. This allows the user to 'refresh' their data or continue downloading if a rate limit was previously reached. 

There are a few instances in which downloading the issue data with comments does not capture all the issues: 

 1. The Github rest API rate limit (currently 5000/hour) is reached. 
 2. Function ends before completing. 
 3. More issues were since posted after downloading. 

This function relies on the naming convention the downloaders utilize on the file to perform this operation. For details, see the function documentation.

```{r Collect all issues x, eval = FALSE}
# Acceptable formats for issue_or_pr
# "is:issue"
# "is:pull-request
issue_or_pr <- "is:issue"
# issue_or_pr <- "is:pull-request"
# gh call but with date
# dir.create(save_path_issue_refresh)
if (issue_or_pr == "is:issue"){
  issue_date_save_path <- save_path_issue_refresh
} else {
  issue_date_save_path <- save_path_pull_request
}
gh_response <- github_api_project_issue_refresh(owner,
                                                repo,
                                                token,
                                                issue_date_save_path,
                                                issue_or_pr,
                                                verbose=TRUE)
github_api_iterate_pages(token,gh_response,
                         issue_date_save_path,
                         prefix="issue",
                         verbose=TRUE)
```
```{r}
# Read all JSON files from the directory
all_issue_files <- lapply(list.files(save_path_issue_refresh, full.names = TRUE), read_json)

# Parse each JSON file using the refresh parser
all_issue <- lapply(all_issue_files, github_parse_search_issues_refresh)


# Combine all the data tables
all_issues_combined <- rbindlist(all_issue, fill = TRUE)

# Display the head of the combined issues table
head(all_issues_combined) %>%
  gt(auto_align = FALSE)
```

The Refresh parser also parses issue data but the folder 'refresh_issues', parsing data retrieved from the search endpoint. In this notebook, that is data from the REFRESH issues portion saved to issue_search directory. 

# Comments

## Issues and Pull Requests Comments

Sometimes it is helpful to retrieve issue comments after a particular date. Issue comments have the drawback in the Github API of only being able to specify a lower bound and downloading all comments created or updated after this date inclusive. If there are multiple revisions of a comment, the most recent version of the comment is downloaded. The acceptable date formats are:

```{r Collect all issue and pull request comments, eval = FALSE}
gh_response <- github_api_project_issue_or_pr_comments(owner,repo,token)

* YYYY-MM-DD
Example: 2023-04-15 (15th April 2023)

* YYYY-MM-DDTHH:MM
Example: 2023-04-15T13:45 (15th April 2023 at 13:45 UTC)

* YYYY-MM-DDTHH:MM:SS
Example: 2023-04-15T13:45:30 (15th April 2023 at 13:45:30 UTC)

* YYYY-MM-DDTHH:MM:SSZ or YYYY-MM-DDTHH:MM:SS+00:00
Example: 2023-04-15T13:45:30Z or 2023-04-15T13:45:30+00:00 (15th April 2023 at 13:45:30 UTC)

```{r Collect comments by date, eval = FALSE}
updated_lower_bound_comment <- "2024-04-25"

# make initial API CALL
gh_response <- github_api_project_issue_or_pr_comments_by_date(owner = owner,
                                                repo = repo,
                                                token = token,
                                                since = updated_lower_bound_comment,
                                                verbose=TRUE)

# Make subsequent API calls and write to JSON file along save path
github_api_iterate_pages(token,gh_response,
                         save_path_issue_or_pr_comments,
                         prefix="issue",
                         verbose=TRUE)
```

## Obtaining author's name and e-mail

The three endpoints used above do not contain author and e-mail information, only the developers GitHub ids. This is a problem, if the project being studied contains communication data outside the GitHub ecosystem (e.g. uses JIRA as issue tracker). In order to link developers to other sources, we need both author and e-mail information.

To do so, we can use the committer endpoint. 

```{r Collect all authors and committers name and e-mail, eval = FALSE}
gh_response <- github_api_project_commits(owner,repo,token)
dir.create(save_path_commit)
github_api_iterate_pages(token,gh_response,
                         save_path_commit,
                         prefix="commit")
```

# Parsing Raw Data to Csv

To parse raw data, we use the associated endpoint parser functions. Keep in mind these functions only parse a subset of all the information in the json ("column wise"). Please consult with the GitHub API or inspect the raw data directly to see all information which is available. 

Note the parsed data will include the **body** column, but it is not shown in the Notebook as it breaks the table formatting due to commas in the text when rendered in HTML.

## Issues 

The issue endpoint provides issue-related metadata such as labels. This can be helpful for the analysis of bug counts, if the project uses bug labels.

```{r}
all_issue <- lapply(list.files(save_path_issue,
                                     full.names = TRUE),read_json)
all_issue <- lapply(all_issue,
                                   github_parse_project_issue)
all_issue <- rbindlist(all_issue,fill=TRUE)

head(all_issue)  %>%
  gt(auto_align = FALSE) 
```

## Pull Requests

Similarly to the Issue API, we can also obtain other metadata from pull requests, including their labels.

```{r}
all_pr <- lapply(list.files(save_path_pull_request,
                                     full.names = TRUE),read_json)
all_pr <- lapply(all_pr,
                                   github_parse_project_pull_request)
all_pr <- rbindlist(all_pr,fill=TRUE)

tail(all_pr,1)  %>%
  gt(auto_align = FALSE) 
```

## Issue or PR Comments

```{r}
all_issue_or_pr_comments <- lapply(list.files(save_path_issue_or_pr_comments,
                                     full.names = TRUE),read_json)
all_issue_or_pr_comments <- lapply(all_issue_or_pr_comments,
                                   github_parse_project_issue_or_pr_comments)
all_issue_or_pr_comments <- rbindlist(all_issue_or_pr_comments,fill=TRUE)

head(all_issue_or_pr_comments,2)  %>%
  gt(auto_align = FALSE) 
```

## Discussion Posts

In order to download data from Github Discussions, we will need to use the Github GraphQL API instead of the REST API, which was used in the previous downloaders. Documentation on GraphQL can be found here: [About the GraphQL API](https://docs.github.com/en/graphql/overview/about-the-graphql-api)

The Github GraphQL API only has 1 endpoint. For downloading data, we only use the queries operation of that endpoint.
To form queries, you need to specify the data you want, which fecthes only the data requested, nothing more.
If information is insufficient, it is advised to check with the API documentation to see what additional information is available. 
More information about what can be retrieved from discussions is located here: [Using GraphQL API for Discussions](https://docs.github.com/en/graphql/guides/using-the-graphql-api-for-discussions)

As stated previously, we download the discussions by forming a query, which only obtains the data requested.
The function by default iterates through all available discussion pages (if user's available tokens allow).
The user can define the maximum number of pages to download in the function parameter (e.g. max_pages=1)

```{r Collect Github Discussions, eval = FALSE, warning=FALSE}
gh_response <- github_api_discussions_refresh(token, owner, repo, save_path_discussions)
```

Two separate tables can be created from the api response.
The first table contains information regarding the initial discussion post.
The second table contains information on the replies and comments for all discussion posts.

```{r Tabulate the results of the Query}
# Parse the discussions
all_discussions <- lapply(list.files(save_path_discussions, full.names = TRUE), read_json)
all_discussions <- lapply(all_discussions, github_parse_discussions)
all_discussions <- rbindlist(all_discussions, fill = TRUE)

# Parse the discussion comments
all_discussion_comments <- lapply(list.files(save_path_discussions, full.names = TRUE), read_json)
all_discussion_comments <- lapply(all_discussion_comments, github_parse_discussion_comments)
all_discussion_comments <- rbindlist(all_discussion_comments, fill = TRUE)

# Display the last 2 rows of each table
tail(all_discussions,2)  %>%
  gt(auto_align = FALSE)
tail(all_discussion_comments,2)  %>%
  gt(auto_align = FALSE)
```

## Refresh Issue or PR Comment

Similar to the refresh of the issues, this chunk allows for the downloading of comments that have been created and/or updated since the most recently created date among data already downloaded. This allows us to 'refresh' the comments, downloading comments made or updated since that date or continue downloading if a rate limit was reached. 

This function relies on the naming convention the downloaders utilize on the file to perform this operation. For details, see the function documentation.

Because the endpoint this function relies on is based on the updated timestamp, running the refresher will download the most recent version of the comment changes. Only the most recent version of the comment will be downloaded, not all copies. However, if the same comment was modified before the next refresh call, then if the refresher function was executed again, then this would result in two comments with the same comment id being present in the table. This can be addressed by performing a group by over the comment\_id in the generated parsed table, and selecting to return the max(updated_at) comment, resulting in a table that only the most recent comment verson as of the latest time the refresher was executed.

```{r Collect all issues comments, eval = FALSE}
#gh call but with date
# get the data
gh_response_issue_or_pr_comment <- github_api_project_issue_or_pr_comment_refresh(owner,                                                                            repo,
                                              token,                                                                            save_path_issue_or_pr_comments,                                                    verbose=TRUE)

# create directory and iterate over data
#dir.create(save_path_issue_or_pr_comments)
github_api_iterate_pages(token,gh_response_issue_or_pr_comment,
                         save_path_issue_or_pr_comments,
                         prefix="issue_or_pr_comment",
                         verbose=TRUE)
```

```{r}
all_issue_or_pr_comments <- lapply(list.files(save_path_issue_or_pr_comments,
                                     full.names = TRUE),read_json)
all_issue_or_pr_comments <- lapply(all_issue_or_pr_comments,
                                   github_parse_project_issue_or_pr_comments)
all_issue_or_pr_comments <- rbindlist(all_issue_or_pr_comments,fill=TRUE)

head(all_issue_or_pr_comments,2)  %>%
  gt(auto_align = FALSE) 

```

# Other API Endpoints

The following API endpoints do not currently implement the refresher, but may provide useful information. 

## Commits Endpoint: Obtaining author's name and e-mail

The three endpoints used before do not contain author and e-mail information, only the developers GitHub ids. This is a problem, if the project being studied contains communication data outside the GitHub ecosystem (e.g. uses JIRA as issue tracker). In order to link developers to other sources, we need both author and e-mail information.

To do so, we can use the committer endpoint. 


```{r Collect all authors and committers name and e-mail, eval = FALSE}
gh_response <- github_api_project_commits(owner,repo,token)
dir.create(save_path_commit)
github_api_iterate_pages(token,gh_response,
                         save_path_commit,
                         prefix="commit",
                         verbose=TRUE)

```

## Issue or PR Endpoint

The refresh search endpoint does **not** include pull requests. If the intent is to download both issues and pull requests, then the `/issue/` endpoint should be used instead:


```{r Collect all issues from issue endpoint, eval = FALSE}
gh_response <- github_api_project_issue(owner,repo,token)
dir.create(save_path_issue)
github_api_iterate_pages(token,gh_response,
                         save_path_issue,
                         prefix="issue")
```

```{r}
all_issue_or_pr <- lapply(list.files(save_path_issue,
                                     full.names = TRUE),read_json)
all_issue_or_pr <- lapply(all_issue_or_pr,
                                   github_parse_project_issue_or_pr_comments)
all_issue_or_pr <- rbindlist(all_issue_or_pr,fill=TRUE)

head(all_issue_or_pr,2)  %>%
  gt(auto_align = FALSE) 
```


## Pull Requests Endpoint

Similarly to the Issue API, we can also obtain other metadata from pull requests, including their labels.

```{r eval = FALSE}
save_path_pull_request <- paste0(save_path,"/pull_request/")
```

```{r Collect all pull requests, eval = FALSE}
gh_response <- github_api_project_pull_request(owner,repo,token)
dir.create(save_path_pull_request)
github_api_iterate_pages(token,gh_response,
                         save_path_pull_request,
                         prefix="pull_request")
```

```{r eval = FALSE}
all_pr <- lapply(list.files(save_path_pull_request,
                                     full.names = TRUE),read_json)
all_pr <- lapply(all_pr,
                                   github_parse_project_pull_request)
all_pr <- rbindlist(all_pr,fill=TRUE)

tail(all_pr,1)  %>%
  gt(auto_align = FALSE) 
```



## Combining Issue and Pull Request communication

If our interest is to observe all the development communication, we may regard both the opening issue, pull request and comments as simply "replies" in a single table. 

Note because we obtain the authors and committers name and e-mail, **only comments made by developers who made at least one commit will contain their name and e-mail**. That is, people who only post issues and comment will not have that information available, and instead will have their github user as part of the `reply_from` column. Therefore, identity match is likely not to work when no author name and e-mail is available. If you are analyzing social smells, this will not be a problem for org silo and missing link (as their condition require code changes). However, since radio silence only consider the mailing list network, caution must be exercised. 


Below we show the result of such merge, including the name and e-mail fields obtained from the commit table. As before, we do not display the body column to prevent breaking the HTML format. 

```{r}
replies <- parse_github_replies(save_path_issue,
                                save_path_pull_request,
                                save_path_issue_or_pr_comments,
                                save_path_commit)  

tail(replies,2)  %>%
  gt(auto_align = FALSE) 
```



