---
title: "Download Mod Mbox and Pipermail Mailing List Archives"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Download Mod Mbox Mailing List Archives}
  %\VignetteEncoding{UTF-8}
---


```{r eval=FALSE}
rm(list = ls())
seed <- 1
set.seed(seed)

# Load libraries
  require(kaiaulu)
  require(data.table)
  require(yaml)
  require(stringi)
  require(XML)
  require(httr)
  require(gt)
```


# Introduction

Open source projects require a means for developers to communicate. These may include mailing lists, issue trackers, discord, etc. This notebooks showcases how to download data from mailing list archives. Two often used archive types are [mod_mbox](https://httpd.apache.org/mod_mbox/) and [pipermail](https://en.wikipedia.org/wiki/GNU_Mailman#cite_note-9), which Kaiaulu offer functions to download data from. The former is commonly used by the Apache Software Foundation projects. The latter, is more commonly use in GNU related projects, but this can vary. 

Each mailing list maintains archives of past messages, often organized by month and year. These archives can be accessed and downloaded for analysis. However, it is important to note that mailing list archives may be split into multiple formats or locations, and not all archives contain the same information. Different archives can differ in completeness, date ranges, and the data they contain. Some archives might lack important fields like "In-Reply-To," which is important for reconstructing message threads. It is, therefore, important the archive being used is carefully selected, since this effects the quality and completeness of analysis.

# Mailing List Organization

Mailing list data is stored in a variety of archives. See:
- Mod Mbox: [Apache Geronimo](https://geronimo.apache.org/mailing-lists.html)).
- Pipermail: [OpenSSL](https://mta.openssl.org/mailman/listinfo/).
(More information on this in the sections below.) This notebook demonstrates how to download and refresh mailing list archives from Mod Mbox and Pipermail.

Mailing lists are typically organized by topic or purpose. For example, the [OpenSSL project](https://www.openssl.org/community/mailinglists.html) maintains several mailing lists, each serving a different group:

- **project-announce**: For important announcements.
- **project-commits**: For commit messages.
- **project-project**: For project discussions.
- **project-users**: For general user questions and discussions.

Mod Mbox archives also organize mailing lists by topic. The apache mailing list archives can be found at https://lists.apache.org/.

# Project Configuration File 

Mailing List archives are hosted by their respective open source projects. Therefore, in order to use Kaiaulu downloaders to obtain mail data, you will need to access the respective open source project, and find out the URL tied to the archive you are interested in. Generally, that is the developer mailing list, if your interest is to understand communication patterns among developers. Alternatively, if the focus of the research is Q&A from the user base, then a user mailing list may make more sense.

Because project lifetime can go as far as a few decades, to have the full picture of what communication took place in the project you may need to download multiple archives and combine them, after turning them into tables using the Kaiaulu parser.

The information you need to find out for each open source project is documented in Kaiaulu using a project configuration file format. For pipermail and mod_mbox this is as follows:

```
mailing_list:
  # for pipermail
  pipermail:
    project_key_1:
      mailing_list: https://mta.openssl.org/pipermail/openssl-users/
      save_folder_path: ../../rawdata/helix/pipermail/save_mbox_mail
      # mbox_file_path is for use only with parse_mbox() function. It is the file to parse
      mbox_file_path: ../../rawdata/helix/pipermail/save_mbox_mail/kaiaulu_202407.mbox
  # for mod mbox
  mod_mbox:
    project_key_1:
      mailing_list: https://lists.apache.org/list.html?announce@apache.org
      save_folder_path: ../../rawdata/helix/mod_mbox/save_mbox_mail
      # mbox_file_path is for use only with parse_mbox() function. It is the file to parse
      mbox_file_path: ../../rawdata/helix/mod_mbox/save_mbox_mail/kaiaulu_202407.mbox
```

The most time intensive step you will be required is to locate the URL of the mailing list archive you wish for in the project website. This is specified under `mailing_list`. Note for pipermail this URL should point to the page containing links to the monthly archives (e.g. https://mta.openssl.org/pipermail/openssl-users/), not the top-level mailing list page that contains all the different types of archives (e.g. https://mta.openssl.org/mailman/listinfo/).


Note: It is important that the paths specified in save_folder_path and mbox_file_path are accurate and do not conflict between projects.

By organizing the configuration in this way, you can manage multiple projects and mailing lists easily. The notebook reads these parameters and uses them to download and process the archives.

Regardless of which mail archive you choose, the downloaders will store the mail data in monthly files, in a `.mbox` format. This is a simple text file that contains some markings to identify the header of the e-mail containing title, authors, etc. You can open any of the .mbox downloaded files with any text editor. 

## Tools Configuration

In addition to the mailing list configurations, you need to specify the path to the [Perceval](https://github.com/chaoss/grimoirelab-perceval) binary in tools.yml. See the wiki for further details on how to setup third party tools. 

Now, you can load the configurations in your R script or notebook using the following code:

```{r eval=FALSE}
# Load tools configuration
tools <- parse_config("../tools.yml")
parse_perceval_path <- get_tool_project("perceval", tools)

# Load project configuration
conf <- parse_config("../conf/helix.yml")
mbox_file_path <- get_mbox_input_file(conf, "project_key_1")
```

# Downloaders and Refreshers

## Downloaders

With the configurations loaded, we can proceed to download the mailing list archives. The downloaders are responsible for fetching the archives from the specified mailing lists and saving them locally in .mbox format.

### Pipermail Downloader

For Pipermail, we need to specify the project key, which is used to retrieve the configuration parameters for the specific project. The project key is used to identify the project in the configuration file.

Now, we can use the getter functions to retrieve the configuration parameters for the specified project key.

```{r eval=FALSE}
conf <- parse_config("../conf/helix.yml")
pipermail_mailing_list <- get_pipermail_domain(conf, "project_key_1")
pipermail_save_folder_path <- get_pipermail_path(conf, "project_key_1")

# Define the date range
pipermail_start_year_month <- 202310
pipermail_end_year_month <- 202405
```

Note that the date range is not set with a getter. The range for downloads changes often, and should be set manually using the YYYYMM format.

With our configurations loaded, we can proceed to downloading the mailing list archives.

```{r eval=FALSE}
# Download archives
download_pipermail(
  mailing_list = pipermail_mailing_list,
  start_year_month = pipermail_start_year_month,
  end_year_month = pipermail_end_year_month,
  save_folder_path = pipermail_save_folder_path,
  verbose = TRUE
)

```

After running this function, the .mbox files will be saved in the specified directory with filenames like kaiaulu_202310.mbox, kaiaulu_202311.mbox, etc.

### Mod Mbox Downloader

The download_mod_mbox() function downloads Mod Mbox archives from a specified Apache Pony Mail mailing list over a given date range. The download_mod_mbox() function downloads Mod Mbox archives by constructing URLs based on the mailing list and date range, saving them as .mbox files named kaiaulu_YYYYMM.mbox.

Similarly to Pipermail, we need to specify the project key for Mod Mbox. The project key is used to retrieve the configuration parameters for the specific project.

Use the getters to extract the parameters:

```{r eval=FALSE}
conf <- parse_config("../conf/helix.yml")
mbox_mailing_list <- get_mbox_domain(conf, "project_key_1")
mbox_save_folder_path <- get_mbox_path(conf, "project_key_1")

# Define the date range
mbox_start_year_month <- 202310
mbox_end_year_month <- 202405
```


The `start_year_month` and `end_year_month` time range parameters should be set manually, as with Pipermail.


```{r eval=FALSE}
download_mod_mbox(
  mailing_list = mbox_mailing_list,
  start_year_month = mbox_start_year_month,
  end_year_month = mbox_end_year_month,
  save_folder_path = mbox_save_folder_path,
  verbose = TRUE
  )

```

After running the function, it constructs URLs like: https://lists.apache.org/api/mbox.lua?list=announce@apache.org&date=2024-01
and saves the files in the specified folder.

## Refreshers

Mailing lists are dynamic, with new emails being added regularly. If you're conducting ongoing analysis or need the most recent data, it's important to refresh your downloaded archives. Manually re-downloading all archives can be time-consuming and inefficient. The refresher functions automate this process by updating only the necessary parts of your archives, saving time and ensuring data completeness. These functions will update your archives by downloading new messages without re-downloading all existing data.

### Pipermail Refresher

The refresh_pipermail function is designed to keep your local archives up-to-date with the latest messages from the mailing list. Here's how it works:

First, it checks the save_folder_path to see if there are any existing files. If the folder is empty, it means you have not downloaded any archives yet. In this case, the function will download all available archives from your specified start_year_month up to the current month, so that you have a complete dataset to work with.

If there are already files in the directory, the function takes a smart approach to updating them. It identifies the most recent archive file based on the filenames (which include the date, like kaiaulu_202311.mbox). It then deletes this most recent file because new messages might have been added to that month since your last download. After deleting it, the function re-downloads this file along with any newer archives that have been added to the mailing list. This way, you don't have to re-download all the archives.

By operating on the directory specified in save_folder_path, the refresh_pipermail function efficiently updates all relevant files, keeping your local archives current without unnecessary downloads.


```{r eval=FALSE}
# Refresh archives
refresh_pipermail(
  mailing_list = pipermail_mailing_list,
  start_year_month = pipermail_start_year_month,
  save_folder_path = pipermail_save_folder_path,
  verbose = TRUE
)

```

This function will ensure that the most recent archives are always up-to-date by redownloading the current month's archive and adding any new months that have been added to the mailing list.

### Mod Mbox Refresher

The behavior is similar to the Pipermail refresher, and makes sure that your Mod Mbox archives are up-to-date.

```{r eval=FALSE}
refresh_mod_mbox(
  mailing_list = mbox_mailing_list,
  start_year_month = mbox_start_year_month,
  save_folder_path= mbox_save_folder_path,
  verbose = TRUE
)
```


# Parsers

After downloading the mailing list archives as .mbox files, the next step is to parse these files to extract meaningful information for analysis. The parse_mbox() function utilizes the Perceval library to parse .mbox files and convert them into structured data tables. This enables easier manipulation and analysis of mailing list data.

## How parse_mbox() Works

The parse_mbox function makes it easy to transform .mbox files into structured data that you can analyze.

The function uses the Perceval library to process .mbox files. Mailing list archives often have variations in their structure—different email headers, missing fields, or inconsistent formats. The parser is designed to handle these variations, so you do not have to worry about cleaning up the data.

As it processes the files, the parser extracts key details from each email, such as the content, sender, recipients and dates. These elements are crucial for understanding communication patterns and building insights.

Finally, to keep things consistent, the function standardizes the column names in the output. Even if the raw data varies from one archive to another, the resulting table will always have predictable and labeled columns, making it easy to work with.

```{r eval=FALSE}
parsed_mail <- parse_mbox(
  perceval_path = parse_perceval_path,
  mbox_file_path = mbox_file_path
)
```

This will store the parsed data into the parsed_mail variable. You can use the gt package to display the parsed data in a readable format:

```{r eval=FALSE}
# Display the first 10 rows of the parsed data using gt
# Refer to the gt documentation for more options on displaying tables
parsed_mail %>%
  head(10) %>%
  gt()
```

Note: Displaying the entire dataset may not be practical if it's large. Showing a sample provides a glimpse of the structure.
