---
title: "Download Mod Mbox and Pipermail Mailing List Archives"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Download Mod Mbox Mailing List Archives}
  %\VignetteEncoding{UTF-8}
---


```{r eval=FALSE}
rm(list = ls())
seed <- 1
set.seed(seed)

# Load libraries
  require(kaiaulu)
  require(data.table)
  require(yaml)
  require(stringi)
  require(XML)
  require(httr)
  require(gt)
```


# Introduction

Mailing list data is stored in a variety of archives. See:
- Mod Mbox: [Apache Geronimo](https://geronimo.apache.org/mailing-lists.html)).
- Pipermail: [OpenSSL](https://mta.openssl.org/mailman/listinfo/).
(More information on this in the sections below.) This notebook demonstrates how to download and refresh mailing list archives from Mod Mbox and Pipermail.

Mailing lists are typically organized by topic or purpose. For example, the [OpenSSL project](https://www.openssl.org/community/mailinglists.html) maintains several mailing lists, each serving a different group:

- **project-announce**: For important announcements.
- **project-commits**: For commit messages.
- **project-project**: For project discussions.
- **project-users**: For general user questions and discussions.

Mod Mbox archives also organize mailing lists by topic. The apache mailing list archives can be found at https://lists.apache.org/.

Each mailing list maintains archives of past messages, often organized by month and year. These archives can be accessed and downloaded for analysis. However, it is important to note that mailing list archives may be split into multiple formats or locations, and not all archives contain the same information. Different archives can differ in completeness, date ranges, and the data they contain. Some archives might lack important fields like "In-Reply-To," which is important for reconstructing message threads. It is, therefore, important the archive being used is carefully selected, since this effects the quality and completeness of analysis.

# Pipermail

## Project Configuration File 

To start, we load the project configuration file, which contains parameters for downloading the mailing list archives. Instead of hard-coding these values in the notebook, we store them in a project configuration file in YAML format. This makes the parameters easier to manage.
Here is an example of the pipermail mailing list section from the configuration file (conf/helix.yml):

```
# top-level key for mailing list config
mailing_list:
  # for pipermail
  pipermail:
    project_key_1:
      mailing_list: https://mta.openssl.org/pipermail/openssl-users/
      start_year_month: 202310
      end_year_month: 202405
      save_folder_path: "../extdata/save_folder_mail"

```

The configuration file contains the following parameters for each mailing list archive:

- project_key_1: A unique key for the project. There can be multiple projects in both the pipermail and mod mbox sections.
- pipermail/ mod_mbox: Indicates whether the setting are for pipermail or mod mbox. Although the parameters are the same, this helps to differentiate between the two types of mailing list archives.
- mailing_list: The URL of the mailing list archive page. Note that this URL should point to the page containing links to the monthly archives (e.g. https://mta.openssl.org/pipermail/openssl-users/), not the top-level mailing list page that contains all the different types of archives (e.g. https://mta.openssl.org/mailman/listinfo/).
- start_year_month: The starting date for downloading archives (in YYYYMM format).
- end_year_month: The ending date for downloading archives (in YYYYMM format).
- save_folder_path: The local directory where the downloaded archives will be saved (if you run the code in this notebook, the archives will be saved in a folder 'extdata', located in the parent directory of kaiaulu (wherever your kaiaulu folder is kept)).

By organizing the configuration in this way, you can manage multiple projects and mailing lists easily. The notebook reads these parameters and uses them to download and process the archives.

## Pipermail Downloader

The following code reads the configuration parameters for project_key_1 of pipermail:

```{r}
conf <- yaml::read_yaml("../conf/helix.yml")
mailing_list <- conf[["mailing_list"]][["pipermail"]][["project_key_1"]][["mailing_list"]]
start_year_month <- conf[["mailing_list"]][["pipermail"]][["project_key_1"]][["start_year_month"]]
end_year_month <- conf[["mailing_list"]][["pipermail"]][["project_key_1"]][["end_year_month"]]
save_folder_path <- conf[["mailing_list"]][["pipermail"]][["project_key_1"]][["save_folder_path"]]
```

After setting the configurations above, you can download the archives using the download_pipermail() function, which downloads and saves .mbox files to the specified directory (save_folder_path). The .mbox files are named with the format kaiaulu_YYYYMM.mbox, where YYYYMM refers to the year and month of the archive.

```{r eval=FALSE}
# Download archives
download_pipermail(
  mailing_list = mailing_list,
  start_year_month = start_year_month,
  end_year_month = end_year_month,
  save_folder_path = save_folder_path,
  verbose = TRUE
)

```

After running this function, the .mbox files will be saved in the specified directory with filenames like kaiaulu_202310.mbox, kaiaulu_202311.mbox, etc.

## Pipermail Refresher

In some cases, you may want to refresh the archive to ensure the most recent months are up-to-date or to handle updates to the mailing list. The refresh_pipermail() function helps automate this process.

How refresh_pipermail Works
1. Checks if the folder is empty: If the folder is empty, it downloads archives starting from start_year_month to the current month using download_pipermail().
2. Finds the most recent file: If the folder is not empty, the function checks for the most recent month’s file (based on the filename) and deletes it.
3. Redownloads from the most recent month: The function then redownloads the archive from the most recent month up to the current month.
# add warning for files do not exist

```{r eval=FALSE}
# Refresh archives
refresh_pipermail(
  mailing_list = mailing_list,
  start_year_month = start_year_month,
  save_folder_path = save_folder_path,
  verbose = TRUE
)

```

This function will ensure that the most recent archives are always up-to-date by redownloading the current month's archive and adding any new months that have been added to the mailing list.

# Mod Mbox

## Project Configuration File

Like in Pipermail, we load the configuration for Mod Mbox from the YAML file, which includes the mailing list URL, the date range, and the save folder path.

Here's an example of the relevant section in the configuration file (conf/helix.yml):

```
# top-level key for mailing list config
mailing_list:
  # for mod mbox
  mod_mbox:
    project_key_1:
      mailing_list: https://lists.apache.org/list.html?announce@apache.org
      start_year_month: 202310
      end_year_month: 202405
      save_folder_path: "../../extdata/save_mbox_mail"

```

The configuration parameters are the same as the ones explained in the section at the top of this notebook, except that the mailing_list should point to a Mod Mbox mailing list URL.

The following code reads the configuration parameters:

```{r eval=FALSE}
conf <- yaml::read_yaml("../conf/helix.yml")
mailing_list <- conf[["mailing_list"]][["mod_mbox"]][["project_key_1"]][["mailing_list"]]
start_year_month <- conf[["mailing_list"]][["mod_mbox"]][["project_key_1"]][["start_year_month"]]
end_year_month <- conf[["mailing_list"]][["mod_mbox"]][["project_key_1"]][["end_year_month"]]
save_folder_path <- conf[["mailing_list"]][["mod_mbox"]][["project_key_1"]][["save_folder_path"]]
```

- mailing_list: The URL of the Mod Mbox mailing list (e.g., https://lists.apache.org/list.html?announce@apache.org).
- start_year_month: The first month to download (format: YYYYMM).
- end_year_month: The last month to download (format: YYYYMM).
- save_folder_path: The directory where the downloaded .mbox files will be saved.

## Mod Mbox Downloader

The download_mod_mbox() function downloads Mod Mbox archives by constructing URLs based on the mailing list and date range, saving them as .mbox files named kaiaulu_YYYYMM.mbox.

```{r eval=FALSE}
download_mod_mbox(
  mailing_list = mailing_list,
  start_year_month = start_year_month,
  end_year_month = end_year_month,
  save_folder_path = save_folder_path,
  verbose = TRUE
  )

```

After running the function, it constructs URLs like: https://lists.apache.org/api/mbox.lua?list=announce@apache.org&date=2024-01
and saves the files in the specified folder.

## Mod Mbox Refresher

To refresh these archives to ensure that you have the latest messages, you can use the refresh_mod_mbox function. This function works similarly to the Pipermail refresher.

How refresh_mod_mbox Works
1. Checks if the folder is empty and, if so, downloads the archives starting from start_year_month to the current month by calling download_mod_mbox().
2. If the folder contains files, it identifies the most recent one using the YYYYMM found in the filename. This file is deleted, and then redownloaded along with all future months.

```{r eval=FALSE}
refresh_mod_mbox(
  mailing_list = mailing_list,
  start_year_month = start_year_month,
  save_folder_path= save_folder_path,
  verbose = TRUE
)
```

This ensures your archive is up-to-date, accounting for new data that may have been added to the mailing list since the last download.

# Parser

After downloading the mailing list archives as .mbox files, the next step is to parse these files to extract meaningful information for analysis. The parse_mbox() function utilizes the Perceval library to parse .mbox files and convert them into structured data tables. This enables easier manipulation and analysis of mailing list data.

## Mbox Parser

The parse_mbox() function takes an .mbox file and parses it into a structured data.table using the Perceval library. 

For the configuration, make sure you have the correct path to the Perceval library in the conf file.

Here's an example of the relevant section in the tools.yml file:

```
perceval: /usr/local/bin/perceval
```

And in the helix.yml configuration file:

```
mailing_list:
  # for mod mbox
  mod_mbox:
    project_key_1:
      mbox_file_path: "../../extdata/save_mbox_mail.kaiaulu_202310.mbox"
```

perceval: found in tools.yml, this should be set to your local path to the perceval binary (use > which perceval to locate the path).
mbox_file_path: should point to the saved .mbox file that will be parsed. See the mbox_path in the mailing_list sections of helix.yml.

Load the configuration:

```{r eval=FALSE}
tools_config <- yaml::read_yaml("../tools.yml")
parse_perceval_path <- tools_config[["perceval"]]

conf <- yaml::read_yaml("../conf/helix.yml")
mbox_file_path <- conf[["mailing_list"]][["mod_mbox"]][["project_key_1"]][["mbox_file_path"]]

```

Run the parser:

```{r eval=FALSE}
parsed_mail <- parse_mbox(
  perceval_path = parse_perceval_path,
  mbox_file_path = mbox_file_path
)
```

This will store the parsed data into the parsed_mail variable. To view the table, use:

```{r eval=FALSE}
# Display the first 10 rows of the parsed data using gt
# Refer to the gt documentation for more options on displaying tables
parsed_mail %>%
  head(10) %>%
  gt()
```

## Retrieve the Latest Mbox File

We can use the parse_mbox_latest_date() function to identify the most recent .mbox file in the specified folder. This can be useful when you want to automate the parsing of the latest data without manually specifying the file name.

First, make sure that the save_folder_path is correctly set to the directory where your .mbox files are stored.

```{r eval=FALSE}
# Get the latest mbox file
latest_mbox_file <- parse_mbox_latest_date(save_folder_path = save_folder_path)
print(latest_mbox_file)
```
This will output the name of the latest .mbox file based on the YYYYMM pattern in the filename.
We can use this to update mbox_file_path to point to the latest file, and call the parse_mbox() function to parse the latest data.

```{r eval=FALSE}
# Update mbox_file_path to use the latest file
mbox_file_path <- file.path(save_folder_path, latest_mbox_file)
print(mbox_file_path)
```

To parse this file: 

```{r eval=FALSE}
# Parse the latest mbox file
parsed_mail <- parse_mbox(
  perceval_path = parse_perceval_path,
  mbox_file_path = mbox_file_path
)
```

Now, parsed_mail contains the parsed data from the latest .mbox file.

```{r eval=FALSE}
# Display the first 10 rows of parsed_mail using gt
# Refer to the gt documentation for more options on displaying tables
parsed_mail %>%
  head(10) %>%
  gt()
```
