---
title: "Download JIRA Issues and Comments"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Download JIRA Issues and Commentss}
  %\VignetteEncoding{UTF-8}
---


```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE,message=FALSE}
require(kaiaulu)
require(data.table)
#require(JirAgileR)
require(knitr, quietly = T)
require(dplyr, quietly = T)
require(jsonlite)
```

# Introduction

This example is adapted from the JirAgileR package [README.md](https://github.com/matbmeijer/JirAgileR). 

As usual, the first step is to load the project configuration file. 

# Project Configuration File

In this notebook, we will obtain data from the issue tracker JIRA. We will use [Apache's Geronimo open source project](https://geronimo.apache.org). Refer to the `conf/` folder on Kaiaulu's git repository for Geronimo and other project configuration files. It is in this project configuration file we specify where Kaiaulu can find the jira sources from Geronimo. We will use the issue_tracker -> jira fields only. In regards to the "issues" and "issue_comments" fields, these should be set to paths where you want to store the jira data. Then, you can access this jira data later using these same paths.

```{r}
conf <- yaml::read_yaml("../conf/geronimo.yml")
issue_tracker_domain <- conf[["issue_tracker"]][["jira"]][["domain"]]
issue_tracker_project_key <- conf[["issue_tracker"]][["jira"]][["project_key"]]
save_path_issue_tracker_issues <- conf[["issue_tracker"]][["jira"]][["issues"]]
save_path_issue_tracker_issue_comments <- conf[["issue_tracker"]][["jira"]][["issue_comments"]]
your_api_token <- scan("~/.ssh/atlassian_token",what="character",quiet=TRUE)
username <- "ssunoo@hawaii.edu"

```

# Specifying the JIRA credentials

In the project configuration file, we define the domain which JIRA is hosted. The example configuration file uses the domain for Apache Software Foundation. We can use this information to queue the domain to identify all project keys available in its JIRA:

```{r eval = FALSE}
# define a function to store the domain to env_name and then save to environment as KAIAULU_DOMAIN
save_jira_credentials <- function(domain=NULL,
                                  username=NULL,
                                  password=NULL,
                                  verbose=FALSE) {
  if (is.null((domain))){
      stop(call.=FALSE, "No domain specified. Check your issue_tracker_domain is correct")}
  
  if(!is.null(domain)){
    env_name <- sprintf("DOMAIN;%s", domain)
  }
  Sys.setenv("KAIAULU_PAT" = env_name)
}


# Save credentials to pass them only one time
save_jira_credentials(domain = issue_tracker_domain)


# Get full list of projects in domain
apache_jira_projects <- data.table(get_jira_projects()) #%>% 
    #select(key, name) 

```

You can load `apache_jira_projects` for the list of projects available. However, this step is optional.

# Download Issues (Without Comments)

In the project configuration file, we also specify the project configuration key (you can also type it directly for the function call). Here, we used the key `GERONIMO`. You can also manually explore the existing issues from a given project by visiting it [on your browser](https://issues.apache.org/jira/projects/GERONIMO/issues/GERONIMO-734?filter=allopenissues). Please see the function documentation warning for `get_jira_issues` on respecting the number of function calls to the server to avoid being IP blocked. 

The following function will download all the available issues. For more details, see the associated package documentation. This data can be used to obtain the `bug_count` metric. See the associated R notebook `bug_count.Rmd` for details.

We will download the data on `verbose`, and as json (`as.data.frame = FALSE`), as the turning the later appears to cause column names to not match values, ignores the requested fields, and gives inconsistent column names. Kaiaulu implements its own parser for the json file, so you will still be able to obtain a table in the end without the associated issues. 

```{r eval = FALSE}
# Retrieve the issues from a single project - in this case the project QTWB from bugreports.qt.io. See documentation to define which fields to see
json_issues <- get_jira_issues(jql_query = paste0("project='",issue_tracker_project_key,"'"),
                fields = c("summary",
                           "description",
                           "creator",
                           "assignee",
                           "reporter",
                           "issuetype",
                           "status",
                           "resolution",
                           "components",
                           "created",
                           "updated",
                           "resolutiondate"),
                username = "your_username",
                password = "your_api_token",
                verbose=TRUE,
                as.data.frame = FALSE)

jsonlite::write_json(json_issues,save_path_issue_tracker_issues)
```








```{r eval=FALSE}
#' @title Define function to fetch issues from JIRA REST API and save as JSON
#' @description Makes a request to JIRA's latest REST API to retrieve all
#' the necessary information regarding the JIRA issues and downloand
#' them to path designated save_path_issue_tracker_issues
#' @param domain Custom JIRA domain URL 
#' @param jql_query the specific query string to specify criteria for fetching
#' @param fields the list of fields that are downloaded in each issue
#' @param maxResults (optional) the maximum number of results to download
#' defaul is 50
#' @param verbose boolean flag to specify printing operational 
#' messages or not
fetch_and_save_jira_issues <- function(domain, 
                                       username = NULL,
                                       password = NULL,
                                       jql_query, 
                                       fields, 
                                       save_path_issue_tracker_issues,  
                                       maxResults = 50, 
                                       verbose = FALSE) {
  
  # Check if domain starts with https:// and if not, prepends it to ensure
  # secure protocol
  if (!grepl("^https?://", domain)) {
    domain <- paste0("https://", domain)
  }
  
  # If username and password is present, then authenticate
  if(!is.null(username)&&!is.null(password)){
    auth <- httr::authenticate(as.character(username), as.character(password), "basic")
  } else {
    auth <- NULL
  }
  
  # Construct the API endpoint URL. appends /rest/api/latest/search to domain
  url <- paste0(domain, "/rest/api/latest/search")
  
  # Prepare query parameters including jql query, list of fields and maxresults to return
  query_params <- list(jql = jql_query, fields = paste(fields, collapse = ","), maxResults = maxResults)
  
  # Make the API call using httr::GET to the constructed URL using the query parameters above
  response <- httr::GET(url, query = query_params)
  
  # Check for HTTP errors and if there is an error, stops the call
  if (httr::http_error(response)) {
    stop("API request failed: ", httr::http_status(response)$message)
  }
  
  # Parse the response - extracts as text - converts JSON to R object and extracts "issues"
  issues <- jsonlite::fromJSON(httr::content(response, "text", encoding = "UTF-8"), simplifyVector = FALSE)$issues
  
  # Save the issues to the specified path
  jsonlite::write_json(issues, save_path_issue_tracker_issues)
  
  # Verbose output
  if (verbose) {
    message("Fetched and saved ", length(issues), " issues to '", save_path_issue_tracker_issues, "'.")
  }
  
  # Returns nothing
  return(invisible(NULL)) # Return invisibly since the primary output is the saved file
}

```

```{r eval=FALSE}
# Define the parameters
# issue_tracker_domain <- conf[["issue_tracker"]][["jira"]][["domain"]]
jql_query <- paste0("project='",issue_tracker_project_key,"'")
fields <- c("summary",
                           "description",
                           "creator",
                           "assignee",
                           "reporter",
                           "issuetype",
                           "status",
                           "resolution",
                           "components",
                           "created",
                           "updated",
                           "resolutiondate")
# save_path_issue_tracker_issues <- conf[["issue_tracker"]][["jira"]][["issues"]]
# save_path_issue_tracker_issue_comments <- conf[["issue_tracker"]][["jira"]][["issue_comments"]]
maxResults <- 100  # Optional: Specify the maximum number of results to fetch
verbose <- TRUE    # Optional: Print operation messages


# Call the function
fetch_and_save_jira_issues(issue_tracker_domain,
                           username, 
                           your_api_token, 
                           jql_query, 
                           fields, 
                           save_path_issue_tracker_issues, 
                           maxResults, 
                           verbose)

```

The json will be downloaded on the path specified in the project configuration file, which by default is `kaiaulu/kaiaulu/rawdata/issue_tracker`. We can then use Kaiaulu's function to parse the data into a tabular format. Since our request did not include the  `comment` field, only the issues table will be available. A few rows of the json issues is shown next:

```{r}
require(kaiaulu)
jira_issues_list <- parse_jira(save_path_issue_tracker_issues)
jira_issues <- jira_issues_list[["issues"]]
jira_comments <- jira_issues_list[["comments"]]
kable(jira_issues[7:8])
```

# Download Issue with Comments

In the same manner as before, we can perform the same function call, but including the field `comment`. This will result in the same table being generated but with the additional comment information per issue (if an issue has more than one comment, the issue id is repeated for each different comment). The comment is shown on the column `comment_comments_id`. 

The data of this table can be used to calculate `social smell metrics`, as it represents a form of developer communication. A notebook discussing how to use JIRA data as communication network and/or combining to mailing list data will be made available in the future. 

```{r eval = FALSE}
json_issue_comments <- get_jira_issues(jql_query = paste0("project='",issue_tracker_project_key,"'"),
                fields = c("summary",
                           "description",
                           "creator",
                           "assignee",
                           "reporter",
                           "issuetype",
                           "status",
                           "resolution",
                           "components",
                           "created",
                           "updated",
                           "resolutiondate",
                           "comment"),
                verbose=TRUE,
                as.data.frame = FALSE)
jsonlite::write_json(json_issue_comments,save_path_issue_tracker_issue_comments)
```

Since this time around we requested the issue data and comments, when using the `parse_jira` function, both the issues and comments table will be available from the parser. Since the issue table was already displayed, the following show a few rows of the issue comments table:

```{r}
jira_issue_comments <- parse_jira(save_path_issue_tracker_issue_comments)
jira_issues <- jira_issue_comments[["issues"]]
jira_comments <- jira_issue_comments[["comments"]]

kable(jira_comments[55:56])
```
