---
title: "Download JIRA Issues and Comments"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Download JIRA Issues and Commentss}
  %\VignetteEncoding{UTF-8}
---


```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE,message=FALSE}
require(kaiaulu)
require(data.table)
require(knitr, quietly = T)
require(dplyr, quietly = T)
require(jsonlite)
```

# Introduction

This example is adapted from the JirAgileR package [README.md](https://github.com/matbmeijer/JirAgileR). 

As usual, the first step is to load the project configuration file. 

# Project Configuration File

In this notebook, we will obtain data from the issue tracker JIRA. We will use [Apache's Geronimo open source project](https://geronimo.apache.org). Refer to the `conf/` folder on Kaiaulu's git repository for Geronimo and other project configuration files. It is in this project configuration file we specify where Kaiaulu can find the jira sources from Geronimo. We will use the issue_tracker -> jira fields only. In regards to the "issues" and "issue_comments" fields, these should be set to paths where you want to store the jira data. Then, you can access this jira data later using these same paths.

# your_api_token is your atlassian token that is used as your password. Your username is your atlassian username (usually your email). These can be created by navigating to the indicated directories and creating a plain text file (vim atlassian_username or vim atlassian_token) and typing in your username or token. Details on how to obtain an atlassian token can be found here: https://support.atlassian.com/atlassian-account/docs/manage-api-tokens-for-your-atlassian-account/

```{r}
conf <- yaml::read_yaml("../conf/geronimo.yml")
issue_tracker_domain <- conf[["issue_tracker"]][["jira"]][["domain"]]
issue_tracker_project_key <- conf[["issue_tracker"]][["jira"]][["project_key"]]
save_path_issue_tracker_issues <- conf[["issue_tracker"]][["jira"]][["issues"]]
save_path_issue_tracker_issue_comments <- conf[["issue_tracker"]][["jira"]][["issue_comments"]]
your_api_token <- scan("~/.ssh/atlassian_token",what="character",quiet=TRUE)
username <- scan("~/atlassian_username",what="character",quiet=TRUE)

```

# function that gets the most recent "created" field of the issues downloaded in you path. This value is used to download only issues created after this date
```{r eval=FALSE}
find_most_recent_issue_date <- function(save_path) {
  files <- list.files(save_path, pattern = "\\.json$", full.names = TRUE)
  most_recent_date <- as.POSIXct("1970-01-01", tz="")  # Initialize with a very old date, no timeszone
  
  for (file in files) {
    message("Scanning file: ", basename(file))
    issue <- jsonlite::read_json(file, simplifyVector = FALSE)
    created_date <- as.POSIXct(issue$created, format="%Y/%m/%d %H:%M", tz="")
    
    if (created_date > most_recent_date) {
      most_recent_date <- created_date
    }
  }
  
  return(most_recent_date)
}

print(find_most_recent_issue_date(save_path_issue_tracker_issues))
```

# Download Issue with Comments

In the same manner as before, we can perform the same function call, but including the field `comment`. This will result in the same table being generated but with the additional comment information per issue (if an issue has more than one comment, the issue id is repeated for each different comment). The comment is shown on the column `comment_comments_id`. 

The data of this table can be used to calculate `social smell metrics`, as it represents a form of developer communication. A notebook discussing how to use JIRA data as communication network and/or combining to mailing list data will be made available in the future. 

Since this time around we requested the issue data and comments, when using the `parse_jira` function, both the issues and comments table will be available from the parser. Since the issue table was already displayed, the following show a few rows of the issue comments table:

# Call the function fetch_and_save_jira_issues (defined on line 55)
```{r eval=FALSE}
# Call the function
all_issues_comments <- download_and_save_jira_issues(issue_tracker_domain,
                           username, 
                           your_api_token, 
                           jql_query = paste0("project='",issue_tracker_project_key,"'"), 
                           fields = c("summary",
                           "description",
                           "creator",
                           "assignee",
                           "reporter",
                           "issuetype",
                           "status",
                           "resolution",
                           "components",
                           "created",
                           "updated",
                           "resolutiondate",
                           "comment"), 
                           save_path_issue_tracker_issue_comments, 
                           maxResults = 5, 
                           verbose = TRUE,
                           maxDownloadsPerHour = 4500)
#path <- file.path(save_path_issue_tracker_issue_comments)
jsonlite::write_json(all_issues_comments, save_path_issue_tracker_issue_comments)
```

#Don't run this because the parser is yet to be built
```{r eval=FALSE}
jira_issue_comments <- parse_jira(save_path_issue_tracker_issue_comments)
jira_issues <- jira_issue_comments[["issues"]]
jira_comments <- jira_issue_comments[["comments"]]
kable(jira_comments[10:11])
```




