---
title: "GitHub API Showcase"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{GitHub API Showcase}
  %\VignetteEncoding{UTF-8}
---

# Introduction

Kaiaulu interface to GitHub API heavily relies on [gh](https://github.com/r-lib/gh), a minimalistic client to access GitHub’s REST and GraphQL APIs. In essence, Kaiaulu only defines a few API endpoints of interest where the tool is currently used, and parses the returned JSON output into a table keeping only fields of interest. But more can be added later. Please see Kaiaulu's Docs Function API to see what is currently available. 

In this Vignette, I will show how to replicate [Aleksander Konnerup data acquisition pipeline](https://github.com/AleksanderKonnerup/AleksanderKonnerup_akon223_projectZ).

## Create a Personal Token 

GitHub limits the number of API calls per IP to only 60 attempts **every hour** at the time this vignette was created. You can check the current rates at [its website](https://docs.github.com/en/free-pro-team@latest/rest/overview/resources-in-the-rest-api#rate-limiting).

If using a personal account token from a free GitHub account, the number of API calls per hour increases to 5000 **per hour**. Therefore, it is recommended you create a personal token by following the [GitHub Documentation instructions](https://docs.github.com/en/free-pro-team@latest/github/authenticating-to-github/creating-a-personal-access-token#:~:text=Creating%20a%20token.%201%20Verify%20your%20email%20address%2C,able%20to%20see%20the%20token%20again.%20More%20items). The process should not take more than 2 minutes.

The GraphQL API will require a token, and does not allow unauthenticated requests. Fortunately both the Github's REST API and GraphQL API can use the same personal account token, given the token has the needed permissions.

The functions in Kaiaulu will assume you have a token available, which can be passed as parameter. 

```{r warning=FALSE,message=FALSE}
rm(list = ls())
require(kaiaulu)
require(data.table)
require(jsonlite)
require(knitr)
require(gh)
```


# GitHub Feature: Assign Issue to Issue Commenters

The feature of interest in this notebook is the introduced capability for project owners to assign issue to issue commenters (i.e. those without merge access), as discussed in the following blog post: https://github.blog/2019-06-25-assign-issues-to-issue-commenters. 

The goal of the following steps is to obtain the data when a project started assigning issues to issue commenters without merge access. 

# Necessary Parameters

To use the pipeline, you must specify the organization and project of interest, and your token.   

```{r}
conf <- parse_config("../conf/kaiaulu.yml")
owner <- get_github_owner(conf, "project_key_1") # Has to match github organization (e.g. github.com/sailuh)
repo <- get_github_repo(conf, "project_key_1") # Has to match github repository (e.g. github.com/sailuh/perceive)
save_path_issue_or_pr_comments <- path.expand(get_github_issue_or_pr_comment_path(conf, "project_key_1"))
save_path_issue_event <- get_github_issue_event_path(conf, "project_key_1")
save_path_commit <- get_github_commit_path(conf, "project_key_1")
save_path_discussions <- get_github_discussions_path(conf, "project_key_1")
# your file github_token contains the GitHub token API obtained in the steps above
token <- scan("~/.ssh/github_token",what="character",quiet=TRUE)
```

# Collecting Data via GitHub API

In this section we obtain the raw data (.json) containing all information the GitHub API endpoint provides. We parse the information of interest in the subsequent section.
Each of the downloaders use 2 functions.
The first function accesses the Github API endpoint using the gh tool.
However, there are limitations to how much information can be downloaded at once. (Typical default is 100 per page)
Therefore the second function saves the information retrieved from the first function to a (.json) file and iterates the page to repeat until all information can be retrieved from the endpoint.

## Using the Github REST API

The Github REST API has many different endpoints available to use, and each retrieves differing levels of information.
Unfortunately, these endpoints also retrieve information that is not of interest, heavily necessitating the parsers to only parse a subset of the information retrieved.

### Issue Events

First we obtain all issue events of the project, so we may later subset issue assignments. 

```{r Collect all issue events, eval = FALSE}
gh_response <- github_api_project_issue_events(owner,repo,token)
github_api_iterate_pages(token,gh_response,save_path_issue_event,prefix="issue_event")
```

### Commits

Next we download commit data from GitHub API. This will be used to know which users in the issue events have or not merge permissions.

```{r Collect all project commit messages, eval = FALSE}
gh_response <- github_api_project_commits(owner,repo,token)
github_api_iterate_pages(token,gh_response,save_path_commit,prefix="commit")
```

## Using the Github GraphQL API

The Github GraphQL API only has 1 endpoint. For downloading data, we only use the queries operation of that endpoint.
To form queries, you need to specify the data you want, which fecthes only the data requested, nothing more.
If information is insufficient, it is advised to check with the API documentation to see what additional information is available.

### Discussions

We download the discussions by forming a query. The query only obtains the data requested.
The function by default iterates through all available discussion pages (if user's available tokens allow).
The user can define the maximum number of pages to download in the function parameter (e.g. max_pages=1)

```{r Collect Github Discussions, eval = FALSE}
gh_response <- github_api_discussions(token, owner, repo, save_path_discussions)
```

# Parsing Raw Data to Csv

To parse raw data, we use the associated endpoint parser functions. Keep in mind these functions only parse a subset of all the information in the json ("column wise"). Please consult with the GitHub API or inspect the raw data directly to see all information that is available. 

```{r Parse the Github Issue Events}
all_issue_event <- lapply(list.files(save_path_issue_event,full.names = TRUE),read_json)
all_issue_event <- lapply(all_issue_event,github_parse_project_issue_events)
all_issue_event <- rbindlist(all_issue_event,fill=TRUE)
all_issue_event[,issue_body:=NULL] # removing column just to prevent html rendering error
kable(head(all_issue_event))
```

```{r Parse the Github Commits}
all_commit <- lapply(list.files(save_path_commit,full.names = TRUE),read_json)
all_commit <- lapply(all_commit,github_parse_project_commits)
all_commit <- rbindlist(all_commit,fill=TRUE)
all_commit[,commit_message:=NULL] # removing column just to prevent html rendering error

kable(head(all_commit))
```

```{r Parse the Github Discussions}
# Parse the discussions
all_discussions <- lapply(list.files(save_path_discussions, full.names = TRUE), read_json)
all_discussions <- lapply(all_discussions, github_parse_discussions)
all_discussions <- rbindlist(all_discussions, fill = TRUE)
all_discussions[,all_discussions:=NULL] # removing column just to prevent html rendering error

# Parse the discussion comments
all_discussion_comments <- lapply(list.files(save_path_discussions, full.names = TRUE), read_json)
all_discussion_comments <- lapply(all_discussion_comments, github_parse_discussion_comments)
all_discussion_comments <- rbindlist(all_discussion_comments, fill = TRUE)
all_discussions[,all_discussion_comments:=NULL] # removing column just to prevent html rendering error

View(all_discussions)
View(all_discussion_comments)
```

# Obtaining Issue Assignments from Non-Committers

With the two tables above, the list of all issue events are calculated and shown below.

```{r}
all_issue_event_assigned <- all_issue_event[event == "assigned"]
assigned_users_without_merge_access <-which(!(all_issue_event_assigned$issue_assignee_login %in%
                                                unique(all_commit$committer_login)))

kable(all_issue_event_assigned[assigned_users_without_merge_access])
```
