---
title: "Issue, Vulnerability and Weakness Networks"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Issue, Vulnerability and Weakness Networks}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE,message=FALSE}
require(kaiaulu)
require(visNetwork)
require(reactable)
require(data.table)
require(igraph)
require(yaml)
```

Load config file.

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/chromium.yml")
perceval_path <- get_tool_project("perceval", tool)
git_repo_path <- get_git_repo_path(conf)
nvdfeed_folder_path <- get_nvdfeed_folder_path(conf)
cveid_regex <- get_cveid_regex(conf)

# Filters
file_extensions <- get_file_extensions(conf)
substring_filepath <- get_substring_filepath(conf)
```

"parse_gitlog" reads the entire commit history of the repository and uses the provided regular expression to extract CVE IDs from the commit messages in the local repo.

A CVE (Common Vulnerabilities and Exposures) identifier is a standardized ID for a publicly known cybersecurity vulnerability in software or hardware. It is used to quickly identify, track, and share information about specific security threats across different databases, tools, and vendors to facilitate patching. The format is CVE-YYYY-NNNNNNN.

```{r}
project_git <- parse_gitlog(perceval_path,git_repo_path,perl_regex = cveid_regex)
#project_git <- parse_gitlog(perceval_path,git_repo_path,save_path)
#project_git <- readRDS(save_path)
```

Filter files based on the conf yml.

```{r}
project_git <- project_git  %>%
  filter_by_file_extension(file_extensions,"file_pathname")  %>% 
  filter_by_filepath_substring(substring_filepath,"file_pathname")
```

Creates a copy of the filtered table for us to work with.

```{r}
project_git_slice <- project_git
```

# Issue Networks

First, we can check the coverage of commits that contain CVE-ID labels. The formula is: (Total unique commits mentioning a CVE ID) / (Total unique commits in the repo) = Percentage of commits referencing a vulnerability (coverage).

```{r}
coverage <- commit_message_id_coverage(project_git,cveid_regex)/length(unique(project_git$data.commit))
coverage
```

```{r}
project_commit_message_id_edgelist <- transform_commit_message_id_to_network(project_git,commit_message_id_regex = cveid_regex)

project_commit_message_id_network <- igraph::graph_from_data_frame(d=project_commit_message_id_edgelist[["edgelist"]], 
                      directed = TRUE, 
                      vertices = project_commit_message_id_edgelist[["nodes"]])
visIgraph(project_commit_message_id_network,randomSeed = 1)
```

If the chosen commit message label is associated to CVEs, one additional metadata of interest to CVEs are their weaknesses. Weaknesses groups related CVE ids. We can obtain the mapping from CVE to CWEs from [NVD Feeds](https://nvd.nist.gov/vuln/data-feeds)

```{r}
# This can take a few minutes!
cve_cwe_mapping <- parse_nvdfeed(nvdfeed_folder_path)
```

The Common Weakness Enumeration (CWE) is a community-driven, standardized dictionary of software and hardware security weaknesses used to identify, classify, and prevent vulnerabilities. It acts as a way to understand root causes of flaws, such as buffer overflows or injection issues, before code is deployed. The format is typically CWE-ID (e.g., CWE-79 or CWE-119).

```{r}
cve_cwe_file_network <- transform_cve_cwe_file_to_network(project_commit_message_id_edgelist,cve_cwe_mapping)
```

```{r}
project_commit_message_id_network <- igraph::graph_from_data_frame(d=cve_cwe_file_network[["edgelist"]], 
                      directed = TRUE, 
                      vertices = cve_cwe_file_network[["nodes"]])
visIgraph(project_commit_message_id_network,randomSeed = 1)
```

