---
title: "Issue, Vulnerability and Weakness Networks"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Issue, Vulnerability and Weakness Networks}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE,message=FALSE}
require(kaiaulu)
require(visNetwork)
require(reactable)
require(data.table)
require(igraph)
require(yaml)
```

Load config file.

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/chromium.yml")
perceval_path <- get_tool_project("perceval", tool)
git_repo_path <- get_git_repo_path(conf)
nvdfeed_folder_path <- get_nvdfeed_folder_path(conf)
cveid_regex <- get_cveid_regex(conf)

# Filters
file_extensions <- get_file_extensions(conf)
substring_filepath <- get_substring_filepath(conf)
```

A CVE (Common Vulnerabilities and Exposures) identifier is a standardized ID for a publicly known cybersecurity vulnerability in software or hardware. It is used to quickly identify, track, and share information about specific security threats across different databases, tools, and vendors to facilitate patching. Using the format is CVE-YYYY-NNNNNNN, we can find CVEs in a given repository. In addition, we can graph this data to visualize which CVEs affect multiple files (Clusters), which files connect to many CVE nodes (Hubs), and which files connect separate clusters (Bridges).

Using the repository's commit history and the regular expression from the conf.yml file to extract CVE IDs from the local commit messages.

```{r}
project_git <- parse_gitlog(perceval_path,git_repo_path,perl_regex = cveid_regex)
#project_git <- parse_gitlog(perceval_path,git_repo_path,save_path)
#project_git <- readRDS(save_path)
```

Filter files based on the conf.yml to remove irrelevant files (such as .txt, .md, and .html), drops paths containing the word "test", and excludes specific parts of the repository.

```{r}
project_git <- project_git  %>%
  filter_by_file_extension(file_extensions,"file_pathname")  %>% 
  filter_by_filepath_substring(substring_filepath,"file_pathname")
```

Copy of the original table with the filters above applied for downstream analysis.

```{r}
project_git_slice <- project_git
```

# Issue Networks

The "coverage" is the percentage of unique commits in the repository that contain a CVE-ID over the total number of unique CVE-IDs. The purpose is to get a high-level overview of a repository's vulnerabilities.

```{r}
coverage <- commit_message_id_coverage(project_git,cveid_regex)/length(unique(project_git$data.commit))
coverage
```

```{r}
project_commit_message_id_edgelist <- transform_commit_message_id_to_network(project_git,commit_message_id_regex = cveid_regex)

project_commit_message_id_network <- igraph::graph_from_data_frame(d=project_commit_message_id_edgelist[["edgelist"]], 
                      directed = TRUE, 
                      vertices = project_commit_message_id_edgelist[["nodes"]])
visIgraph(project_commit_message_id_network,randomSeed = 1)
```

If the chosen commit message label is associated to CVEs, one additional metadata of interest to CVEs are their weaknesses. Weaknesses group related CVE ids. We can obtain the mapping from CVE to CWEs from [NVD Feeds](https://nvd.nist.gov/vuln/data-feeds)

```{r}
# This can take a few minutes!
cve_cwe_mapping <- parse_nvdfeed(nvdfeed_folder_path)
```

The Common Weakness Enumeration (CWE) is a community-driven, standardized dictionary of software and hardware security weaknesses used to identify, classify, and prevent vulnerabilities. A "weakness" is a condition in software, firmware, hardware, or a service that, under certain conditions, can lead to a vulnerability (https://cwe.mitre.org/about/index.html). The format is typically CWE-ID (e.g., CWE-79 or CWE-119).

```{r}
cve_cwe_file_network <- transform_cve_cwe_file_to_network(project_commit_message_id_edgelist,cve_cwe_mapping)
```

The resulting graph treats CVEs and CWEs as being in a similar group to one another. This allows us to map the clusters, hubs, and bridges not just by specific exposures, but by the underlying architectural weaknesses.

```{r}
project_commit_message_id_network <- igraph::graph_from_data_frame(d=cve_cwe_file_network[["edgelist"]], 
                      directed = TRUE, 
                      vertices = cve_cwe_file_network[["nodes"]])
visIgraph(project_commit_message_id_network,randomSeed = 1)
```

