---
title: "Collaboration Network Parser with Vulnerabilities"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Collaboration Network Parser with Vulnerabilities}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE,message=FALSE}
require(kaiaulu)
require(visNetwork)
require(reactable)
require(data.table)
require(igraph)
require(yaml)
```

Load config file.

```{r}
tool <- yaml::read_yaml("../tools.yml")
conf <- yaml::read_yaml("../conf/chromium.yml")
perceval_path <- tool[["perceval"]]
git_repo_path <- conf[["data_path"]][["git"]]
nvdfeed_folder_path <- conf[["data_path"]][["nvd_feed"]]
cveid_regex <- conf[["commit_message_id_regex"]][["cve_id"]]

# Filters
file_extensions <- conf[["filter"]][["keep_filepaths_ending_with"]]
substring_filepath <- conf[["filter"]][["remove_filepaths_containing"]]
```

# Construct Contribution Network

```{r}
#save_path <- "/Users/cvp/Desktop/kaiaulu_project_input/rds/APR_git.rds"
project_git <- parse_gitlog(perceval_path,git_repo_path,perl_regex = cveid_regex)
#project_git <- parse_gitlog(perceval_path,git_repo_path,save_path)
#project_git <- readRDS(save_path)
```

# Filter files

```{r}
project_git <- project_git  %>%
  filter_by_file_extension(file_extensions,"file")  %>% 
  filter_by_filepath_substring(substring_filepath,"file")
```


```{r}
project_git_slice <- project_git

project_contribution_network <- parse_gitlog_network(project_git_slice,mode="author")
project_contribution_network <- igraph::graph_from_data_frame(d=project_contribution_network[["edgelist"]], 
                      directed = TRUE, 
                      vertices = project_contribution_network[["nodes"]])
visIgraph(project_contribution_network,randomSeed = 1)
```


# Construct a Commit-File Network

```{r}
project_commit_network <- parse_gitlog_network(project_git_slice,mode="commit")
project_commit_network <- igraph::graph_from_data_frame(d=project_commit_network[["edgelist"]], 
                      directed = TRUE, 
                      vertices = project_commit_network[["nodes"]])
visIgraph(project_commit_network,randomSeed = 1)
```

# Co-Change Network

```{r}
co_change_network <- igraph::bipartite_projection(project_commit_network,
                                          multiplicity = TRUE,
                                          which = FALSE) # FALSE is the file projection
visIgraph(co_change_network,randomSeed = 1)
```

## Issue Networks

First, we can check the coverage of commits which contain cve-id labels

```{r}
coverage <- commit_message_id_coverage(project_git,cveid_regex)/length(unique(project_git$data.commit))
coverage
```


```{r}
project_commit_message_id_edgelist <- parse_commit_message_id_network(project_git,commit_message_id_regex = cveid_regex)

project_commit_message_id_network <- igraph::graph_from_data_frame(d=project_commit_message_id_edgelist[["edgelist"]], 
                      directed = TRUE, 
                      vertices = project_commit_message_id_edgelist[["nodes"]])
visIgraph(project_commit_message_id_network,randomSeed = 1)
```
If the chosen commit message label is associated to CVEs, one additional metadata of interest to CVEs are their weaknesses. Weaknesses groups related CVE ids. We can obtain the mapping from CVE to CWEs from [NVD Feeds](https://nvd.nist.gov/vuln/data-feeds)

```{r}
# This can take a few minutes!
cve_cwe_mapping <- parse_nvdfeed(nvdfeed_folder_path)
cve_cwe_file_network <- parse_cve_cwe_file_network(project_commit_message_id_edgelist,cve_cwe_mapping)
```

```{r}
project_commit_message_id_network <- igraph::graph_from_data_frame(d=cve_cwe_file_network[["edgelist"]], 
                      directed = TRUE, 
                      vertices = cve_cwe_file_network[["nodes"]])
visIgraph(project_commit_message_id_network,randomSeed = 1)
```



```{r eval = FALSE,include=FALSE}
## Co-Change adjacency matrix 
reactable(as_adjacency_matrix(co_change_network,attr = "weight",sparse = F))
```
