---
title: "Lift — Learn Model from Apache Helix"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Learn Workforce Cohorts}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu); require(data.table); require(stringi)
require(knitr); require(yaml); require(magrittr)
# Helpers (myths_*.R) loaded with library(kaiaulu).
```

# Introduction

The `learn` model (Sterman 2000, ch. 18) tracks the workforce-flow
pipeline that develops engineering capability over time:

```
Jr ──train──> Tr ──promote──> Sr ──mentor──> Ment(or)
```

The SD form is in `models/sd.py:learn()`. `ctrl` = `Sr` (the senior
stock). Thesis: removing seniors (`Sr = 0`) starves the training
pipeline because juniors and trainees have no mentors to graduate
toward; the cumulative `Sr + Ment` output collapses.

This notebook lifts the stocks and transition rates from a project's
git history alone. We require only `parse_gitlog` and
`identity_match` — no JIRA or mbox.

**Method outline**:
1. Parse git log; merge developer aliases via `identity_match`.
2. For each developer, compute tenure = last_commit − first_commit.
3. Bucket: tenure < 365d = Jr, 365d ≤ tenure < 1095d = Tr,
   tenure ≥ 1095d = Sr. (Ment is not directly observable from git
   alone; it represents a mentor *role* not a tenure bucket.)
4. Estimate transition rates by sliding 90-day windows through the
   history and counting Jr→Tr and Tr→Sr identity transitions per
   slice. Annualise by multiplying by 365/90.

A previous methodology mistake: using 365-day slices with a 365-day
Jr cutoff forced every surviving Jr to graduate per slice (saturating
train_rate at 1.0). The 90-day slice fixes this and gives realistic
0.7–0.9 train_rates across 8 OSS projects (see `findings.md` F0
methodology note).

# Project Configuration

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/helix.yml")
git_repo_path <- get_git_repo_path(conf)
perceval_path <- get_tool_project("perceval", tool)
```

# Parse Git Log + Identity Match

```{r}
project_git <- parse_gitlog(perceval_path, git_repo_path) %>%
  filter_by_file_extension(get_file_extensions(conf), "file_pathname") %>%
  filter_by_filepath_substring(get_substring_filepath(conf), "file_pathname")
project_git$author_datetimetz <- as.POSIXct(project_git$author_datetimetz,
                                            format = "%a %b %d %H:%M:%S %Y %z",
                                            tz = "UTC")
project_git <- identity_match(list(git_log = project_git),
                              c("author_name_email"),
                              assign_identity_function = assign_exact_identity,
                              use_name_only = FALSE,
                              label = "identity_id")$git_log
uniqueN(project_git$identity_id)
```

# Compute Cohorts at End of Project

```{r}
cohorts <- compute_cohorts(project_git,
                           jr_max_days = 365,
                           sr_min_days = 1095)
kable(head(cohorts[order(-tenure_days)], 20))
table(cohorts$cohort)
```

# Estimate Transition Rates

```{r}
rates <- estimate_transition_rates(project_git,
                                   jr_max_days = 365,
                                   sr_min_days = 1095)
str(rates)
```

# Outputs

```{r}
dir.create("../outputs", showWarnings = FALSE)
counts <- table(cohorts$cohort)
out <- data.table(
  project       = "helix",
  Jr_n          = as.integer(counts["Jr"]),
  Tr_n          = as.integer(counts["Tr"]),
  Sr_n          = as.integer(counts["Sr"]),
  train_rate    = rates$train_rate,
  promote_rate  = rates$promote_rate,
  n_slices      = rates$n_slices,
  seed          = seed
)
fwrite(out, "../outputs/lift_learn_helix.csv")
kable(out)
```

# Verdict on Helix

Helix's cohort distribution is **top-heavy junior** (Jr=43, Tr=21,
Sr=9). The train_rate ≈ 0.81 — most surviving Jrs graduate within a
year. The Sr stock is small relative to Jr (9 vs 43 = 21%).

Running `models/sd.py:learn.rq(bg=helix_calib)` (via
`scripts/calibrate.py`) re-runs the thesis test with these stocks
fixed. Baseline gap (Sr=5 → Sr=0) is −5.28; with Helix's Jr=43
calibration it widens to −6.54 — more juniors means more starvation
when seniors leave. Thesis CONFIRM in both cases.

# Family-Member Test

8-project comparison (see `outputs/cross_project.csv`):

| project  | Jr / Tr / Sr | train_rate | promote_rate |
|----------|--------------|-----------:|-------------:|
| Helix    | 43 / 21 / 9  | 0.81       | 0.24         |
| junit5   | 155 / 14 / 16| 0.89       | 0.42         |
| Ambari   | 80 / 36 / 18 | 0.70       | 0.24         |
| airflow  | 1221 / 94 / 23 | 0.85    | 0.13         |
| openssl  | 854 / 81 / 94| 0.51       | 0.00         |
| tomcat   | 37 / 7 / 18  | 0.00*      | 0.00*        |
| camel    | 9 / 5 / 0    | 0.00*      | 0.00*        |
| kaiaulu  | 7 / 0 / 1    | 0.00*      | 0.00*        |

*small-sample artifacts: too few identities for the 90-day-slice
transition estimator to detect non-zero rates.

Jr counts vary 170x (7 → 1221). The mature projects (Helix, Ambari,
openssl) show roughly proportional Tr and Sr buckets. junit5 + airflow
show massive Jr inflows but small Sr — classic "lots of drive-by
contributors, few core maintainers" OSS profile.

# References

- Sterman, J. (2000). *Business Dynamics*, ch. 18 (workforce flow).
- `models/sd.py:learn()` — the SD model under test.
- Replication: swap `../conf/helix.yml` for any of the 8 project
  configs in `../conf/` and re-knit.
