---
title: "Download GitHub Pull Request Comments via REST API"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Download GitHub Pull Request Comments via REST API}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

In this vignette, the goal is to download all comments on the GitHub
Pull Requests Conversations tab. In a GitHub Pull Request, comments
written in the conversations tab are not easily retrievable by one
endpoint in chronological order. They are instead split into four
mutually exclusive endpoints. The four REST API endpoints being used are
pull requests, issue/pr comments, pull request in-line comments, and
pull request reviews.

-   Pull Requests

    -   Reviews

        -   In-Line PR Comment

    -   Pull Request Comments

## Libraries

To use the notebook, you will need to ensure you have the required R
packages installed.

```{r warning=FALSE,message=FALSE}
rm(list = ls())
require(kaiaulu)
require(data.table)
require(jsonlite)
require(knitr)
require(magrittr)
require(gt)
```

## Project Configuration File

To use the pipeline, you must specify the organization and project of
interest.

```{r warning=FALSE}
conf <- parse_config("../conf/kaiaulu.yml")
owner <- get_github_owner(conf, "project_key_1") # Has to match github organization (e.g. github.com/sailuh)
repo <- get_github_repo(conf, "project_key_1") # Has to match github repository (e.g. github.com/sailuh/perceive)

# Path you wish to save all raw data.
save_path_issue_refresh <- get_github_issue_search_path(conf, "project_key_1")
save_path_issue <- get_github_issue_path(conf, "project_key_1")
save_path_pull_request <- get_github_pull_request_path(conf, "project_key_1")
save_path_pr_comments <- get_github_pr_comments_path(conf, "project_key_1")
save_path_issue_or_pr_comments <- get_github_issue_or_pr_comment_path(conf, "project_key_1")
save_path_commit <- get_github_commit_path(conf, "project_key_1")
save_path_pr_reviews <- get_github_pr_review_path(conf, "project_key_1")
save_path_pr_commits <- get_github_pr_commits_path(conf, "project_key_1")
save_path_pr_files <- get_github_pr_files_path(conf, "project_key_1")
save_path_pr_reviewers <- get_github_pr_reviewers_path(conf, "project_key_1")
save_path_pr_merge <- get_github_pr_merge_path(conf, "project_key_1")
# Create all folder directories
#create_file_directory(conf)
```

## Create a Personal Token

GitHub limits the number of API calls per IP to only 60 attempts **every
hour** at the time this vignette was created. You can check the current
rates at [its
website](https://docs.github.com/en/free-pro-team@latest/rest/overview/resources-in-the-rest-api#rate-limiting).

If using a personal account token from a free GitHub account, the number
of API calls per hour increases to 5000 **per hour**. Therefore, it is
recommended you create a personal token by following the [GitHub
Documentation
instructions](https://docs.github.com/en/free-pro-team@latest/github/authenticating-to-github/creating-a-personal-access-token#:~:text=Creating%20a%20token.%201%20Verify%20your%20email%20address%2C,able%20to%20see%20the%20token%20again.%20More%20items).
The process should not take more than 2 minutes.

The token should then be copied and pasted into `../.ssh/github_token`.
Both the `.ssh` folder and `github_token` file may need to be created
manually. Check for hidden folders before creating the `.ssh` folder.
The `github_token` file should not have any file extensions.

```{r Scan GitHub Token}
# your file github_token (a text file) contains the GitHub token API
token <- scan("~/.ssh/github_token",what="character",quiet=TRUE)
```

## GitHub Pull Request Comments

In GitHub, a Pull Request conversation starts off with the Pull Request
description. Pull Request comments are made by users. A reviewer can
start a review where they can add a single comment, request changes, or
approve for merging. This starts the review comment. From there, in-line
code comments can be made from the review.

# Download GitHub Pull Request Comments

## Conversation Comments

### Pull Request Comments

Pull Request Comments not connected to a review, appear in the GitHub
Pull Request Conversations tab but is retrieved by the issue comments
endpoint. Using this endpoint, both issues and pull request comments are
retrieved.

```{r Collect comments by pull request number}
#gh call but with date
# get the data
gh_response_issue_or_pr_comment <- github_api_project_issue_or_pr_comment_refresh(owner,repo,token,save_path_issue_or_pr_comments,verbose=TRUE)

# create directory and iterate over data
#dir.create(save_path_issue_or_pr_comments)
github_api_iterate_pages(token,gh_response_issue_or_pr_comment,
                         save_path_issue_or_pr_comments,
                         prefix="issue_or_pr_comment",
                         verbose=TRUE)
```

It is important to filter out the issue comments in this notebook. The
key to this is in the html_url of each comment, differentiated by
/issues or /pull

```{r}
pr_comments <- lapply(list.files(save_path_issue_or_pr_comments, full.names = TRUE),read_json)
pr_comments <- lapply(pr_comments, github_parse_project_issue_or_pr_comments)
pr_comments <- rbindlist(pr_comments,fill=TRUE)

#filter only pull requests using grepl
pr_comments <- pr_comments[grepl("/pull/", html_url)]

head(pr_comments,10)  %>%
  gt(auto_align = FALSE) 
```

### First Pull Request Comment

We may notice the first comment of the pull request is missing. The pull
request description is treated as a pull request comment in the
conversations tab by GitHub. However, it is not retrievable by the other
comments endpoints. To get this, we must utilize the pull request
endpoint to gather information on the pull request, and retrieve the
body paragraph for that pull request.

```{r Collect information on pull request}
issue_or_pr <- "is:pull-request"

gh_response <- github_api_project_issue_refresh(owner, repo, token, save_path_pull_request, issue_or_pr)
github_api_iterate_pages(token, gh_response, save_path_pull_request, prefix="pull-request", verbose=TRUE)
```

```{r}
pull_requests <- lapply(list.files(save_path_pull_request,
                                     full.names = TRUE),read_json)
pull_requests <- lapply(pull_requests,
                                   github_parse_search_issues_refresh)
pull_requests <- rbindlist(pull_requests,fill=TRUE)

head(pull_requests,10)  %>%
  gt(auto_align = FALSE) 
```

## Code Comments

### Pull Request Reviews

Up until now, previous comments were not associated to code review.
Users can perform code reviews by add all in-line code review comments
and save them together with "Changes Requested" or "Approved" or just a
single comment in-line rather than grouping. The review endpoint allows
analysts to see the state of the review such as requesting changes,
approving, or a singular comment. With this the review has its own
comment body different from the in-line code comments. More information
on the fields can be seen on the function definition for
`github_parse_project_pr_reviews`. These fields of data can greatly
contribute to finding and analyzing patterns in collaboration between
developers.

```{r Collect Reviews from Pull Requests}
gh_response <- github_api_pr_reviews_refresh(owner, repo, token, save_path_pull_request, save_path_pr_reviews)
```

```{r Parse Review Comments from Pull Requests}
pr_reviews <- lapply(list.files(save_path_pr_reviews, full.names = TRUE), read_json)
pr_reviews <- lapply(pr_reviews, github_parse_project_pr_reviews)
pr_reviews <- rbindlist(pr_reviews, fill = TRUE)
head(pr_reviews,10)  %>%
  gt(auto_align = FALSE) 
```

### Pull Request In-Line Code Comments

Other information we can obtain is the in-line code comments from Pull
Requests. The comments include the diff_hunk field, which is a string
that contains the in-line code reviewers may comment on the pull
request. The particular line numbers the comment is referencing is
specified by the review comment. These in-line code comments have a
correlating review id number that matches the review endpoint and other
in-line code comments, therefore grouping them all together. More
information on the fields can be seen on the function definition for
`github_parse_project_pr_comments`. These fields of data can greatly
contribute to finding and analyzing patterns in collaboration between
developers.

```{r Collect Comments from Pull Requests}
gh_response <- github_api_project_pr_comments_refresh(owner, repo, token)
github_api_iterate_pages(token, gh_response, save_path_pr_comments, prefix="pr_comments")
```

```{r Parse Comments from Pull Requests}
inline_comments <- lapply(list.files(save_path_pr_comments, full.names = TRUE), read_json)
inline_comments <- lapply(inline_comments, github_parse_project_pr_comments)
inline_comments <- rbindlist(inline_comments, fill = TRUE)
head(inline_comments,10)  %>%
  gt(auto_align = FALSE) 
```

# Pull Request Conversations Tab

After understanding the four endpoints that make up GitHub
Conversations, we want to visualize that in the Kaiaulu. Here we gather
specified data from pull_requests, pr_comments, pr_reviews, and in-line
comments and the goal is to organize them by PR Review number and time.

```{r Tabulate Endpoints}
pull_requests <- lapply(list.files(save_path_pull_request,full.names = TRUE),read_json)
pull_requests <- lapply(pull_requests, github_parse_search_issues_refresh)
pull_requests <- rbindlist(pull_requests,fill=TRUE)
pull_requests <- pull_requests[,.(html_url=html_url, created_at=created_at, body=body )]

pr_comments <- lapply(list.files(save_path_issue_or_pr_comments, full.names = TRUE), read_json)
pr_comments <- lapply(pr_comments, github_parse_project_pr_comments)
pr_comments <- rbindlist(pr_comments, fill = TRUE)
pr_comments <- pr_comments[grepl("/pull/", html_url)]
pr_comments <- pr_comments[,.(html_url=html_url, created_at=created_at, body=body )]

inline_comments <- lapply(list.files(save_path_pr_comments, full.names = TRUE), read_json)
inline_comments <- lapply(inline_comments, github_parse_project_pr_comments)
inline_comments <- rbindlist(inline_comments, fill = TRUE)
inline_comments <- inline_comments[,.(html_url=html_url, PR_review=review_id, created_at=created_at, body=body )]

pr_reviews <- lapply(list.files(save_path_pr_reviews, full.names = TRUE), read_json)
pr_reviews <- lapply(pr_reviews, github_parse_project_pr_reviews)
pr_reviews <- rbindlist(pr_reviews, fill = TRUE)
pr_reviews <- pr_reviews[,.(html_url=html_url, PR_review=review_id, created_at=submitted_at, body=body, state=state)]

pr_endpoints <- rbind(pull_requests, pr_comments, inline_comments, pr_reviews, fill = TRUE)

pr_endpoints[, thread_root := ifelse(is.na(PR_review), 
                                    NA,
                                    html_url)]

# Step 1: Add a sort key that gives priority to review summary (state not NA)
pr_endpoints[, is_review_summary := !is.na(state)]

# Step 2: Use PR_review as a grouping variable — fallback to NA if no review ID
pr_endpoints[, thread_id := fcoalesce(as.character(PR_review), html_url)]

# Step 3: Order by thread ID, then review summary (TRUE = 1, sorted before FALSE = 0), then timestamp
pr_endpoints <- pr_endpoints[order(thread_id,
                                   -as.integer(is_review_summary),
                                   created_at)]

head(pr_endpoints,100)  %>%
  gt(auto_align = FALSE) 
```
