---
title: "Radio Silence Showcase"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Radio Silence Showcase}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE,message=FALSE}
require(kaiaulu)
require(igraph)
require(visNetwork)
require(data.table)
require(yaml)
require(stringi)
require(knitr)
```

# Introduction

This notebook demonstrates how to compute radio silence — a social smell that identifies developers who act as sole bridges between otherwise disconnected communication communities. The pipeline uses the OpenSSL project's pipermail mailing list data.

Radio silence boundary spanners are authors who are the only link between two distinct developer communities. If they stop communicating, the communities lose contact entirely, creating a coordination risk.

The pipeline proceeds as follows:

1. Parse pipermail mbox files into a reply table
2. Transform into a bipartite author-thread network
3. Project to a unimodal author-author network
4. Detect communities with OSLOM
5. Count radio silence boundary spanners with `smell_radio_silence`

# Project Configuration

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/openssl.yml")

oslom_undir_path <- get_tool_project("oslom_undir", tool)
perceval_path    <- get_tool_project("perceval", tool)

mbox_path <- get_pipermail_path(conf, "project_key_1")
```

# Parse Mbox Replies

Parse all pipermail mbox files into a unified reply table. Empty files are skipped.

```{r}
project_mbox <- data.table()
mbox_files <- list.files(mbox_path, full.names = TRUE)
mbox_files <- mbox_files[file.size(mbox_files) > 0]
for (mbox_file_path in mbox_files) {
  project_mbox <- rbind(project_mbox, parse_mbox(perceval_path, mbox_file_path), fill = TRUE)
}

project_mbox$reply_datetimetz <- as.POSIXct(project_mbox$reply_datetimetz,
                                             format = "%a, %d %b %Y %H:%M:%S %z", tz = "UTC")

kable(head(project_mbox[, .(reply_from, reply_subject, reply_datetimetz)]))
```

# Build Bipartite Author-Thread Network

Transform the reply table into a bipartite graph where one node type represents authors and the other represents mail threads (subjects).

```{r}
reply_network <- transform_reply_to_bipartite_network(project_mbox)
kable(head(reply_network[["nodes"]]))
kable(head(reply_network[["edgelist"]]))
```

# Project to Author-Author Network

Project the bipartite graph onto the author node type to obtain a unimodal author-author co-communication network. An edge between two authors means they both posted in the same thread.

```{r}
author_network <- bipartite_graph_projection(reply_network, mode = TRUE,
                                             weight_scheme_function = weight_scheme_sum_edges)
kable(head(author_network[["nodes"]]))
kable(head(author_network[["edgelist"]]))
```

## Visualize Author-Author Network

```{r fig.width=10, fig.height=7}
is_directed <- any(author_network[["edgelist"]][["direction"]] == "directed")
i_author_graph <- igraph::graph_from_data_frame(
  d        = author_network[["edgelist"]],
  directed = is_directed,
  vertices = author_network[["nodes"]]
)
visIgraph(i_author_graph, randomSeed = seed)
```

# Community Detection

Detect communication communities using OSLOM. Each cluster groups authors who communicate densely with one another.

```{r}
mail_clusters <- community_oslom(oslom_undir_path, author_network, seed = seed, n_runs = 1000)
kable(mail_clusters[["assignment"]])
kable(mail_clusters[["info"]])
```

# Radio Silence

Count the number of boundary spanners — authors who are the sole bridge between otherwise disconnected communities. An output of 0 means all communities are either well-connected or isolated with no single bottleneck author.

```{r}
brokers <- smell_radio_silence(mail.graph = author_network, clusters = mail_clusters)
cat("Number of radio silence boundary spanners:", length(brokers), "\n")
if (length(brokers) > 0) {
  kable(data.table(boundary_spanner = brokers))
}
```
