---
title: "Mailing List and Issue Tracker Reply Showcase"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Mailing List and Issue Tracker Reply Showcase}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE,message=FALSE}
require(kaiaulu)
require(visNetwork)
require(reactable)
require(data.table)
require(stringi)
require(igraph)
require(yaml)
```

Load config file.

```{r}
tool <- yaml::read_yaml("../tools.yml")
conf <- yaml::read_yaml("../conf/geronimo.yml")
perceval_path <- tool[["perceval"]]

mbox_path <- conf[["mailing_list"]][["mbox"]]
jira_issue_comments_path <- conf[["issue_tracker"]][["jira"]][["issue_comments"]]
```


Since Jira issue creators and the issue description, and the comment creator and the comment body can be understood as e-mail relies, we can format the issue data into the same reply format as mailing lists:


```{r}
project_mbox <- parse_mbox(perceval_path,mbox_path)
project_jira <- parse_jira_replies(parse_jira(jira_issue_comments_path))
```

# Identity Matching

```{r}
#Identity matching
project_log <- list(project_mbox=project_mbox,project_jira=project_jira)
project_log <- identity_match(project_log,
                                name_column = c("reply_from","reply_from"),
                                assign_identity_function = assign_exact_identity,
                                use_name_only=TRUE,
                                label = "raw_name")
project_mbox <- project_log[["project_mbox"]]
project_jira <- project_log[["project_jira"]]
```

```{r}
intersect(project_mbox$identity_id,project_jira$identity_id)
```

# Datetime Parsing

```{r}
# Parse timestamps and convert to UTC
project_mbox$reply_datetimetz <- as.POSIXct(project_mbox$reply_datetimetz,
                                        format = "%a, %d %b %Y %H:%M:%S %z", tz = "UTC")

project_jira$reply_datetimetz <- as.POSIXct(project_jira$reply_datetimetz,
                                        format = "%Y-%m-%dT%H:%M:%S.000+0000", tz = "UTC")

# Remove comments without timestamp
project_mbox <- project_mbox[!is.na(reply_datetimetz)]
project_jira <- project_jira[!is.na(reply_datetimetz)]


project_mbox_slice <- project_mbox[reply_datetimetz >= as.POSIXct("2005-08-01", format = "%Y-%m-%d",tz = "UTC") & reply_datetimetz < as.POSIXct("2005-08-30", format = "%Y-%m-%d",tz = "UTC")]

project_jira_slice <- project_jira[reply_datetimetz >= as.POSIXct("2005-08-01", format = "%Y-%m-%d",tz = "UTC") & reply_datetimetz < as.POSIXct("2005-08-30", format = "%Y-%m-%d",tz = "UTC")]
```

# Mailing List

## Construct Mailing List Communication Network

```{r}
project_mbox_network <- transform_reply_to_bipartite_network(project_mbox_slice)

project_mbox_network <- igraph::graph_from_data_frame(d=project_mbox_network[["edgelist"]], 
                      directed = TRUE, 
                      vertices = project_mbox_network[["nodes"]])

visIgraph(project_mbox_network,randomSeed = 1)
```

## Construct Issue Communication Network

```{r}
project_jira_network <- transform_reply_to_bipartite_network(project_jira_slice)

project_jira_network[["nodes"]][type == FALSE]$color <- '#3895D3'

project_jira_network <- igraph::graph_from_data_frame(d=project_jira_network[["edgelist"]], 
                      directed = TRUE, 
                      vertices = project_jira_network[["nodes"]])

visIgraph(project_jira_network,randomSeed = 1)
```

# Merging Communication Networks

```{r}
both_mediums_authors <- intersect(project_mbox_slice$identity_id,project_jira_slice$identity_id)
both_mediums_authors
```


Recall each row obtained from `parse_mbox` and `parse_jira_comments` represent a comment from an author. Moreover, because we performed identity match, all comments are identified by a common id. Therefore, all is needed to combine the data is combine them row-wise as we would any other table in R: 

```{r}
project_jira_slice$reply_subject <- stringi::stri_c('#3895D3',project_jira_slice$reply_subject,sep = " ")
```


```{r}
project_communication <- rbind(project_mbox_slice,project_jira_slice)
```

Since the set union of the tables maintain the same format, we can use the same previous code to transform the data in a graph and visualize it: 

```{r}
project_communication_network <- transform_reply_to_bipartite_network(project_communication)

both_mediums_authors <- intersect(project_mbox_slice$raw_name,project_jira_slice$raw_name)


# Distinguish JIRA comments with a slightly darker tone of blue
subject_node_names <- project_communication_network[["nodes"]][type == FALSE]$name
is_jira_node <- stringi::stri_detect_regex(subject_node_names,pattern='#3895D3')
project_communication_network[["nodes"]][type == 'FALSE'][is_jira_node]$color <- '#3895D3'

# Distinguish Authors who use both JIRA and Mailing list in red 
project_communication_network[["nodes"]][type == TRUE & (name %in% both_mediums_authors)]$color <- 'red'

project_communication_network <- igraph::graph_from_data_frame(d=project_communication_network[["edgelist"]], 
                      directed = TRUE, 
                      vertices = project_communication_network[["nodes"]])

visIgraph(project_communication_network,randomSeed = 1)
```

