---
title: "Lift — Rework Model from Apache Helix"
output:
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Rework Szz Cycle}
  %\VignetteEncoding{UTF-8}
---

```{r}
rm(list = ls())
seed <- 1
set.seed(seed)
```

```{r warning=FALSE, message=FALSE}
require(kaiaulu); require(data.table); require(stringi)
require(knitr); require(yaml); require(magrittr)
# Helpers (myths_*.R) loaded with library(kaiaulu).
```

# Introduction

The `rework` model (Abdel-Hamid & Madnick 1991) expresses the hidden
rework cycle: Req → Dev → Test → (pass | Rew → Dev again). Its `ctrl`
parameter is `failrate`. Thesis: `failrate` 0.1 → 0.7 lets rework
dominate, killing net Done output.

This notebook lifts `failrate` from Apache Helix as the share of
commits in each window that the SZZ pass marks as bug-introducing.

# Project Configuration

```{r}
tool <- parse_config("../tools.yml")
conf <- parse_config("../conf/helix.yml")
perceval_path <- get_tool_project("perceval", tool)
git_repo_path <- get_git_repo_path(conf)
```

# Parse Git Log

```{r}
project_git <- parse_gitlog(perceval_path, git_repo_path) %>%
  filter_by_file_extension(get_file_extensions(conf), "file_pathname") %>%
  filter_by_filepath_substring(get_substring_filepath(conf), "file_pathname")
project_git$author_datetimetz <- as.POSIXct(project_git$author_datetimetz,
                                            format = "%a %b %d %H:%M:%S %Y %z",
                                            tz = "UTC")
```

# Load SZZ Pairs

```{r}
szz <- parse_szz_bugfixes("../data/helix/derived/szz_pairs.csv")
nrow(szz)
```

# Compute failrate per Window

```{r}
window_days <- 90
fr <- compute_failrate_per_window(szz, project_git,
                                  window_days = window_days)
kable(head(fr, 20))
summary(fr$failrate)
```

# Verdict Interpretation

`rework`'s thesis predicts net `Done` drops when `failrate` crosses
the 0.5 mark. Compare Helix's median `failrate` to that threshold.

# Outputs

```{r}
dir.create("../outputs", showWarnings = FALSE)
out <- data.table(project        = "helix",
                  failrate_median = median(fr$failrate, na.rm = TRUE),
                  failrate_mean   = mean(fr$failrate,   na.rm = TRUE),
                  n_windows       = nrow(fr),
                  window_days     = window_days,
                  seed            = seed)
fwrite(out, "../outputs/lift_rework_helix.csv")
kable(out)
```
