---
title: "Understand Showcase"
output: 
  html_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Understand Showcase}
  %\VignetteEncoding{UTF-8}
---


# Introduction

Within a project, we might want to see the in-between file or in-between class dependencies. This notebook showcases how to obtain either dependencies using [Scitool's Understand](https://scitools.com)). Note you will need a license from Scitools to use this Notebook.


```{r warning = FALSE, message = FALSE}
rm(list = ls())
require(kaiaulu)
require(visNetwork)
require(XML)
require(stringi)
require(igraph)
require(data.table)
require(gt)
```

# Project Configuration File

For this notebook, we will use Apache Helix as an example. Refer to Helix project configuration file under the conf/ folder, in particular the `understand` section. For supported languages, see [scitools documentation](https://support.scitools.com/support/solutions/articles/70000582794-supported-languages).

```{r}
tool <- parse_config("../tools.yml")
scitools_path <- get_tool_project("scitools", tool)

conf <- parse_config("../conf/helix.yml")
keep_dependencies_type <- get_understand_keep_dependencies_type(conf)
project_path <- get_understand_project_path(conf)

# Scitools
understand_folder <- get_understand_output_path(conf)
code_language <- get_understand_code_language(conf)

db_path <- stringi::stri_c(understand_folder,"Understand.und")

file_dependencies_path <- stringi::stri_c(understand_folder,"file_dependencies.xml")
class_dependencies_path <- stringi::stri_c(understand_folder,"class_dependencies.xml")
```


# Build Understand DB

To export dependencies, scitools require a `und` database to be created:

```{r eval = FALSE}
db_path <- build_understand_project(scitools_path = scitools_path, project_path = project_path, language = code_language, output_dir = understand_folder)
```

# Export Dependencies

Subsequently, we can make our dependencies export out of the database. We can either export XML files or class dependencies.

## File Dependencies

```{r eval = FALSE}
file_dependencies_path <- export_understand_dependencies(scitools_path = scitools_path, db_filepath = db_path, parse_type = "file", output_filepath = file_dependencies_path)
```

The XML files can then be parsed into tables for manipulation in Kaiaulu:

```{r}
file_dependencies <- parse_understand_dependencies(dependencies_path = file_dependencies_path)

head(file_dependencies[["node_list"]])  %>%
  gt(auto_align = FALSE)
```

```{r}
head(file_dependencies[["edge_list"]])  %>%
  gt(auto_align = FALSE)
```


## Class Dependencies

The process to export and parse class dependencies is identical, save for specifying the `parse_type` to "class":

```{r eval = FALSE}
class_dependencies_path <- export_understand_dependencies(scitools_path = scitools_path, db_filepath = db_path, parse_type = "class", output_filepath = class_dependencies_path)
```


```{r}
class_dependencies <- parse_understand_dependencies(dependencies_path = class_dependencies_path)
                                                   
head(class_dependencies[["node_list"]])  %>%
  gt(auto_align = FALSE)
```

```{r}
head(class_dependencies[["edge_list"]])  %>%
  gt(auto_align = FALSE)
```


# Network Visualization

We can display the parsed dependencies as a network. You can use the dependency_kind parameter to subset the dependencies used. Refer to [Kaiaulu wiki](https://github.com/sailuh/kaiaulu/wiki/Scitools) for the types of dependencies Scitools support.

## File Network

```{r}
file_graph <- transform_understand_dependencies_to_network(parsed = file_dependencies, weight_types = c(keep_dependencies_type[2], keep_dependencies_type[5]))


project_function_network <- igraph::graph_from_data_frame(d=file_graph[["edge_list"]], 
                      directed = TRUE,
                      vertices = file_graph[["node_list"]])
visIgraph(project_function_network,randomSeed = 1)
```


## Class Network

The same applies to outputting our class_dependencies, but we can visualize what our class dependency data is using the same parameters sans the parsed data table.

```{r}
class_graph <- transform_understand_dependencies_to_network(parsed = class_dependencies, weight_types = c(keep_dependencies_type[2], keep_dependencies_type[5]))

project_function_network <- igraph::graph_from_data_frame(d=class_graph[["edge_list"]], 
                      directed = TRUE,
                      vertices = class_graph[["node_list"]])
visIgraph(project_function_network,randomSeed = 1)
```

